---
title: "Behavioral Classification"
author: "Ben Weinstein"
date: "May 5, 2016"
output: html_document
---

## Feature descriptors

Following [DeRuiter](http://arxiv.org/pdf/1602.06570v1.pdf)
Using the high-resolution, multivariate tag data and the the visual observations of behaviour and spatial locations, a number of variables were chosen to summarise the whales’ behaviour. These variables were computed on a dive-by-dive basis; in other words, the input data for modelling were
time series for which the sampling unit was one dive. Here, a “dive” was defined as any excursion. from the surface to 10 m depth or greater. Dive start- and end-times were detected in the dive profile by visual inspection. The variables calculated for each dive were dive duration, post-dive
surface duration, maximum depth, step length and turning angle in the horizontal dimension, number of feeding lunges, and variability of heading. Dive duration was the time (in seconds) from
the start of a dive until the first surfacing following the dive. The post-dive surface duration was
defined as the time (in seconds) from the end of one dive until the start of the subsequent dive.
Maximum dive depth was the maximum depth attained by the whale over the course of the dive.
To compute step lengths and turning angles based on position observations collected by human
observers, we first used linear interpolation on the visual observation data on whale positions, which
were unevenly spaced in time. The resulting interpolated tracks had a position estimate at the
mid-point time of each dive

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Code and Output

```{r}
summary(cars)
```

## Slide with Plot

```{r, echo=FALSE}
plot(cars)
```

