---
title: "Antarctic Whale Project: Data Exploration"
author: "Ben Weinstein"
date: "`r Sys.time()`"
output: 
  html_document:
    keep_md: yes
    number_sections: yes
    theme: spacelab
    toc: yes
    pandoc_args: [
    "-M2GB", "+RTS", "-K64m", "-RTS"
    ]
---

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width = 11,fig.height = 5,cache=F)
library(boot)
library(raster)
library(reshape2)
library(ggplot2)
library(survival)
library(MASS)
library(ggmap)
library(dplyr)
library(chron)
library(gridExtra)
library(stringr)
library(R2jags)
library(move)
library(doSNOW)
library(foreach)

#Source model file
source("RunModel.R")
```

```{r,eval=F}
#Load in data
load("DataExploration.RData")
```

```{r}
#read data
mdat<-read.csv("InputData/Antarctic Humpback overlap with krill fisheries -3085052751565071394.csv")

#name env columns
#Create a oceandepth in km 
mdat$ocean<-round(mdat$ETOPO1.Elevation * -1 /1000,3)
mdat$coast<-round(mdat$NASA.Distance.to.Coast,1)
mdat$SST<-mdat$MODIS.Ocean.Terra.OceanColor.4km.Yearly.Daytime.SST
mdat$ChloroA<-log(mdat$MODIS.Ocean.Aqua.OceanColor.4km.Monthly.Chlorophyll.A)
mdat$NPP<-log(mdat$OSU.Ocean.NPP.0.17deg.Monthly.NPP)

#standardize column names to match the simulation
#Create an animal tag.
mxy <- as(mdat, "data.frame")
mxy$Animal<-mxy$individual.local.identifier
mxy$x<-mxy$location.long
mxy$y<-mxy$location.lat

#remove minke whales
minke<-c("131117","131118","131120","154184")
mxy<-mxy[!mxy$individual.local.identifier %in% minke,]

#grab set of animals

mxy<-mxy[mxy$Animal %in% c("112699","121207"),]

#empty coordinates
mxy<-mxy[!is.na(mxy$x),]

#crop by extent
d<-SpatialPointsDataFrame(cbind(mxy$x,mxy$y),data=mxy,proj4string=CRS("+proj=longlat +datum=WGS84"))

d<-crop(d,y=extent(c(-75,-50,-70,-57.5)))
#give it a geospatial layer


mxy<-as.data.frame(d)

#set datestamp
mxy$timestamp<-as.POSIXct(mxy$timestamp,format="%Y-%m-%d %H:%M:%S.000")

#month and year columns
mxy$Month<-months(mxy$timestamp)
mxy$Year<-years(mxy$timestamp)

#Only austral sping and summer
mxy<-mxy[mxy$Month %in% month.name[1:7],]

#remove empty timestamps
mxy<-mxy[!is.na(mxy$timestamp),]

#remove duplicates
mxy<-mxy[!duplicated(data.frame(mxy$timestamp,mxy$Animal)),]

#no distance to coast,elevation
mxy<-mxy[!mxy$ETOPO1.Elevation>0,]
```

```{r}
#hold copy of orig
oxy<-mxy

#remake spatial object
d<-SpatialPointsDataFrame(cbind(mxy$x,mxy$y),data=mxy,proj4string=CRS("+proj=longlat +datum=WGS84"))

#Plot
pp<-c(mean(mxy$x),mean(mxy$y))
m <- get_map(location=pp,source="google",zoom=4,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal)),size=.5) + scale_color_discrete(guide="none") + theme_inset()
```

##By Month

```{r,fig.height=12}
mxy$Month<-factor(mxy$Month,levels=month.name)
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal)),size=.5)  + facet_wrap(~Month) + scale_color_discrete(guide="none")
```

#Correlated random walk

*Process Model*

$$ d_{t} \sim T*d_{t-1} + Normal(0,\Sigma)$$
$$ x_t = x_{t-1} + d_{t} $$

## Parameters

For each individual:

$$\theta = \text{Mean turning angle}$$
$$\gamma = \text{Move persistence} $$

For both behaviors process variance is:
$$ \sigma_{latitude} = 0.1$$
$$ \sigma_{longitude} = 0.1$$

##Behavioral States

$$ \text{For each individual i}$$
$$ Behavior_1 = \text{traveling}$$
$$ Behavior_2 = \text{foraging}$$

$$ \alpha_{i,1,1} = \text{Probability of remaining traveling when traveling}$$
$$\alpha_{i,2,1} = \text{Probability of switching from Foraging to traveling}$$

$$\begin{matrix}
  \alpha_{i,1,1} & 1-\alpha_{i,1,1} \\
  \alpha_{i,2,1} & 1-\alpha_{i,2,1} \\
\end{matrix}
$$

##Environment

Behavioral states are a function of local environmental conditions. The first environmental condition is ocean depth. I then build a function for preferential foraging in shallow waters.

It generally follows the form, conditional on behavior at t -1:

$$Behavior_t \sim Multinomial([\phi_{traveling},\phi_{foraging}])$$

With the probability of switching states:

$$logit(\phi_{traveling}) = \alpha_{Behavior_{t-1}} + \beta_{Month,1} * Environment_{y[t,]}$$

$$logit(\phi_{foraging}) = \alpha_{Behavior_{t-1}} $$

Following Bestley in preferring to describe the switch into feeding, but no estimating the resumption of traveling.

The effect of the environment is temporally variable such that

$$ \beta_{Month,2} \sim ~ Normal(\beta_{\mu},\beta_\tau)$$


##Continious tracks

The transmitter will often go dark for 10 to 12 hours, due to weather, right in the middle of an otherwise good track. The model requires regular intervals to estimate the turning angles and temporal autocorrelation. As a track hits one of these walls, call it the end of a track, and begin a new track once the weather improves. We can remove any micro-tracks that are less than three days.
Specify a duration, calculate the number of tracks and the number of removed points. Iteratively.

```{r}
##Time is the beginning of the first point.
step_length=6

sxy<-split(mxy,mxy$Animal)

#time diff function
timed<-function(d,step_length){
  d$j[1]<-0
  for (x in 2:nrow(d)){
    d$j[x]<-as.numeric(difftime(as.POSIXct(d$timestamp[x]),as.POSIXct(d$timestamp[x-1]),units="mins"))/(step_length*60)
  }
  
  #Split out track endings
  ends<-c(1,which(d$j>1),nrow(d))

  for(w in 2:length(ends)){
    d[ends[w-1]:ends[w],"Track"]<-w-1
  }
  
  #remove tracks that are shorter than three days
  track_time<-d %>% group_by(Track) %>% summarize(mt=difftime(max(as.POSIXct(timestamp)),min(as.POSIXct(timestamp)),units="hours")) %>% filter(mt>=24) %>% .$Track
  
  d<-d[d$Track %in% track_time,]
  
  #renumber the tracks
  d$Track<-as.numeric(as.factor(d$Track))
  return(d)
  }

sxy<-lapply(sxy,timed,step_length=6)

#Format matrices for jags
mxy<-rbind_all(sxy)

######recode whales
#mxy$Animal<-as.numeric(as.factor(mxy$Animal))

sxy<-split(mxy,list(mxy$Animal,mxy$Track),drop=TRUE)

sxy<-lapply(sxy,function(x){
#How many observations in each step length segment
x$Step<-as.numeric(cut(as.POSIXct(x$timestamp),"6 hours"))
return(x)
})

mxy<-rbind_all(sxy)

#refactor animal
mxy$Animal<-as.numeric(as.factor(mxy$Animal))
```

How did the filter change the extent of tracks?

```{r}
#Plot
mxy<-mxy %>% arrange(Animal,Track,Step)

pp<-c(mean(mxy$x),mean(mxy$y))
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal),group=paste(Animal,Track)),size=.5) + scale_color_discrete(guide="none") + theme_inset()
```

#Environmental Models

Looping through each covariate.

* Bathymetry
* Distance to coast
* Sea surface temperature
* Chlorophyl A
* Wave Height
* Primary Productivity
* Sea Ice Cover (%)

```{r}
names_of_vars<-c("SST","ChloroA","Ocean","Coast","NPP")

cl<-makeCluster(3,"SOCK")
registerDoSNOW(cl)
posteriors<-foreach(x=1:length(names_of_vars),.packages=c("dplyr","reshape2","R2jags","stringr")) %dopar% {
  runModel(names_of_vars[[x]],mxy)
}
stopCluster(cl)

#Name by variable
names(posteriors)<-names_of_vars
pc<-melt(posteriors,id.vars=colnames(posteriors[[1]]))
pc$L1<-as.factor(pc$L1)
levels(pc$L1)<-names_of_vars
```


##Chains
```{r,fig.height=20}
ggplot(data=pc,aes(x=Draw,y=value)) + geom_line(aes(col=chain)) + facet_grid(par~L1,scales='free') + theme_bw()  + ggtitle("Estimated parameters")
```


```{r,fig.height=20}
#Plot
ggplot(data=pc,aes(x=value)) + geom_histogram() + facet_grid(par~L1,scales='free') + theme_bw()  + ggtitle("Estimated parameters")
```


## Parameter Summary
```{r,fig.height=10,fig.width=15}
psum<-pc %>% group_by(L1,parameter,par) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))
ggplot(psum,aes(x=par,y=mean,ymax=upper,ymin=lower)) + geom_linerange(aes(col=L1),position=position_dodge(width = 0.1)) + geom_point(aes(col=L1),position=position_dodge(width = 0.1)) + facet_wrap(~parameter,scales="free",ncol=2)
```

## Which estimates do not overlap with zero?

```{r}
psum<-pc %>% group_by(L1,parameter,par)%>% filter(parameter %in% c("beta_mu","beta"))  %>%  summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

sig<-psum %>% group_by(L1,par) %>% summarize(Significant=!findInterval(0,sort(c(lower,upper)))==1)
```

#Behavior and environment

##Hierarchical 

```{r}
#get the posterior estimates of the env function
plotall<-list()
  
  for(j in 1:length(posteriors)){
    plotall[[j]]<-posteriors[[j]] %>% filter(parameter %in% c('beta_mu','alpha_mu')) %>% dcast(.,Draw+chain~par) %>% group_by(Draw,chain) %>% do(dplot(a1=c(.$'alpha_mu[1]',.$'alpha_mu[2]'),beta=c(.$'beta_mu[1]',.$'beta_mu[2]'),x=unique(mxy[[names(posteriors)[j]]]))) %>% group_by(x=x,Begin,State) %>% summarize(mean=mean(y,na.rm=T),lower=quantile(y,0.05,na.rm=T),upper=quantile(y,0.95,na.rm=T))  
  }
names(plotall)<-names_of_vars
plotall<-melt(plotall,id.vars=colnames(plotall[[1]]))

ggplot(plotall) + geom_ribbon(aes(x=x,y=mean,ymin=upper,ymax=lower,fill=State),alpha=0.5) + theme_bw() + labs(col="Transition",y="Probability",fill="Transition") + facet_wrap(~L1,scales="free")

ggplot(plotall) + geom_line(aes(x=x,y=mean,ymin=upper,ymax=lower),alpha=0.5) + theme_bw() + labs(col="Transition",y="Probability",fill="Transition") + facet_wrap(~L1,scales="free")
```

### Zoom in


```{r}

limits<-plotall %>% group_by(L1) %>% summarize(min=quantile(x,0,na.rm=T),max=quantile(x,0.75,na.rm=T))
limits<-limits %>% mutate(L1=factor(L1,levels=names(posteriors))) %>% arrange(L1)
p<-list()
for (x in 1:length(posteriors)){
p[[x]]<-ggplot(plotall[plotall$L1 %in% names(posteriors)[x],]) + geom_line(aes(x=x,y=mean,ymin=upper,ymax=lower),alpha=0.5) + theme_bw() + labs(col="Transition",y="Probability",fill="Transition")  + scale_x_continuous(limits=c(limits$min[x],limits$max[x])) + ggtitle(names(posteriors)[x])
}
p
```


```{r}
pribbon<-list()
for (x in 1:length(posteriors)){
pribbon[[x]]<-ggplot(plotall[plotall$L1 %in% names(posteriors)[x],]) + geom_ribbon(aes(x=x,y=mean,ymin=upper,ymax=lower),alpha=0.5) + theme_bw() + labs(col="Transition",y="Probability",fill="Transition")  + scale_x_continuous(limits=c(limits$min[x],limits$max[x])) + ggtitle(names(posteriors)[x])
}
pribbon
```

## By Month

```{r}

monthall<-list()
  
  for(j in 1:length(posteriors)){
    
  #get the posterior estimates of the env function
  postplot_alpha<-posteriors[[j]] %>% filter(parameter %in% c('alpha_mu')) %>% do(dcast(.,Draw+chain~par))
  postplot_beta<-posteriors[[j]] %>% filter(parameter %in% c('beta')) %>% dcast(.,Month+Draw+chain~parameter+Behavior)
  postall<-merge(postplot_alpha,postplot_beta,by=c("Draw","chain"))

  monthall[[j]]<-postall %>% group_by(Month,Draw,chain) %>% do(dplot(a1=c(.$'alpha_mu[1]',.$'alpha_mu[2]'),beta=c(.$'beta_1',.$'beta_2'),x=unique(mxy[[names(posteriors)[j]]]))) %>% group_by(x=x,Month,Begin,State) %>% summarize(mean=mean(y,na.rm=T),lower=quantile(y,0.05,na.rm=T),upper=quantile(y,0.95,na.rm=T))
   }

names(monthall)<-names_of_vars
monthall<-melt(monthall,id.vars=colnames(monthall[[1]]))

monthall$Month<-factor(monthall$Month,levels=month.name)

ggplot(monthall) + geom_ribbon(aes(x=x,y=mean,ymin=upper,ymax=lower,fill=Month),alpha=0.5) + theme_bw() + labs(col="Transition",y="Probability of foraging when traveling",fill="Transition") + facet_wrap(~L1,scales="free")

ggplot(monthall) + geom_line(aes(x=x,y=mean,ymin=upper,ymax=lower,col=Month)) + theme_bw() + labs(col="Transition",x="Ocean Depth (km)",y="Probability of Switching to Foraging to Traveling",fill="Transition") + facet_wrap(~L1,scales="free")

```

```{r}
#check out the memory sizing
lsos()
#notify end of run
#size of the run
gc()
save.image("DataExploration.RData")
```
