---
title: "moveHMM"
author: "Ben Weinstein"
date: "August 15, 2016"
output: html_document
---

#Format Data

```{r}
```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width = 11,fig.height = 5)
library(boot)
library(raster)
library(reshape2)
library(ggplot2)
library(survival)
library(MASS)
library(ggmap)
library(dplyr)
library(chron)
library(gridExtra)
library(stringr)
library(R2jags)
library(moveHMM)

#New model to be run, flag turned off if just updating.
newModel<-T
```

```{r,eval=F}
#Load in data
load("SingleSpecies.RData")
#jags$recompile()
newModel<-F
```

```{r}
#raster layers
#ice
ice<-raster("InputData/ETOPO1_Ice_g_geotiff.tif")
ice<-round(ice*-1/1000,3)
rice<-aggregate(ice,fact=0.1/res(ice))

#coast
distc<-raster("InputData/DistancetoCoast.tif")
rice[rice<0]<-NA
distc[is.na(rice)]<-NA

#read data
mdat<-read.csv("InputData/Antarctic Humpback overlap with krill fisheries -1133643618235885002.csv")

##need to look into making into a movebank object.

#name env columns
#Create a oceandepth in km 
mdat$ocean<-round(mdat$ETOPO1.Elevation * -1 /1000,3)
mdat$coast<-round(mdat$NASA.Distance.to.Coast,1)
mdat$SST<-mdat$MODIS.Ocean.Aqua.OceanColor.4km.Monthly.Daytime.SST
mdat$ChloroA<-mdat$MODIS.Ocean.Aqua.OceanColor.4km.Monthly.Chlorophyll.A

#standardize column names to match the simulation
#Create an animal tag.
mxy <- as(mdat, "data.frame")
mxy$Animal<-mxy$individual.local.identifier
mxy$x<-mxy$location.long
mxy$y<-mxy$location.lat

#grab set of animals
#mxy<-mxy[mxy$Animal %in% c("112699","121207","131151","112703","121211","131132","112737","131156","12326","123232","112705","154187","121210","131134","131141","112701","131159","131142","131144","154186","123224"),]

#mxy<-mxy[mxy$Animal %in% c("131134","131141","112701","131159","131142","131144","154186","123224"),]
mxy$Animal<-as.numeric(as.factor(mxy$Animal))

#empty coordinates
mxy<-mxy[!is.na(mxy$x),]

#crop by extent
d<-SpatialPointsDataFrame(cbind(mxy$x,mxy$y),data=mxy)
d<-crop(d,y=extent(c(-75,-50,-70,-55)))
mxy<-as.data.frame(d)

#set datestamp
mxy$timestamp<-as.POSIXct(mxy$timestamp,format="%Y-%m-%d %H:%M:%S.000")

#month and year columns
mxy$Month<-months(mxy$timestamp)
mxy$Year<-years(mxy$timestamp)

#remove empty timestamps
mxy<-mxy[!is.na(mxy$timestamp),]

#remove duplicates
mxy<-mxy[!duplicated(data.frame(mxy$timestamp,mxy$Animal)),]
```

#Prep Data

```{r}
mh<-mxy %>% select(x=location.long,y=location.lat,ID=Animal,ocean,coast) %>% filter(ID %in% 1:10)
prep<-prepData(mh,type="LL")
head(prep)
```

#Fit Model

```{r}
## initial parameters for gamma and von Mises distributions
mu0 <- c(0.1,1) # step mean (two parameters: one for each state)
sigma0 <- c(0.1,1) # step SD
zeromass0 <- c(0.1,0.05) # step zero-mass
stepPar0 <- c(mu0,sigma0,zeromass0)
angleMean0 <- c(pi,0) # angle mean
kappa0 <- c(1,1) # angle concentration
anglePar0 <- c(angleMean0,kappa0)

m <- fitHMM(data=prep,nbStates=2,stepPar0=stepPar0,
anglePar0=anglePar0,formula=~ocean)
```

#Plot Model

```{r}
par(ask=FALSE)
devAskNewPage(ask = F)
plot(m)
```