---
title: "Antarctic Whale Project: MultiSpecies Simulation"
author: "Ben Weinstein"
date: "`r Sys.time()`"
output: 
  html_document:
    keep_md: true
    number_sections: true
    toc: true
    theme: spacelab
---

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width = 10,fig.height = 5)
library(boot)
library(raster)
library(reshape2)
library(ggplot2)
library(MASS)
library(dplyr)
library(chron)
library(gridExtra)
library(stringr)
library(R2jags)
library(survival)

#New model to be run, flag turned off if just updating.
newModel<-TRUE

#Source Mountains function, thanks to Sam Urmy http://www.oceanographerschoice.com/2010/10/fractal-landscapes-in-r-part-two/
source("Mountains.R")
```

```{r,eval=F}
#Load in data
load("MultiSpecies.RData")
jags$recompile()
newModel<-F
```

#Abstract
I simulated correlated random walks with similar properties to previous marine pinnepid studies. The virtue of the simulation is that we can build complexity slowly. At each step we can verify that the model captures the true, known, relationship. Once we have developed a model that satisfies our aims, we can then apply it to the observed data.

## Current State

A dynamic correlated random walk with two states (traveling, feeding) that are a function of local environmental conditions. There are multiple individuals whose covariates are drawn from a group level mean. Observation error is modeled as normal distribution. Irregularly spaced observations are indexed to discrete time steps.

## To add
* Observation Error based on location class and long tails (Brost 2015)
* Posterior Model Checks
* Seasonal Variation
** Variation in process error (Breed 2012)
** Variation in diel?

#Correlated random walk

*Process Model*

$$ d_{t} \sim T*d_{t-1} + Normal(0,\Sigma)$$
$$ x_t = x_{t-1} + d_{t} $$

## Parameters

For each individual:

$$\theta = \text{Mean turning angle}$$
$$\gamma = \text{Move persistence} $$

For both behaviors process variance is:
$$ \sigma_{latitude} = 0.1$$
$$ \sigma_{longitude} = 0.1$$

##Behavioral States

$$ \text{For each individual i}$$
$$ Behavior_1 = \text{traveling}$$
$$ Behavior_2 = \text{foraging}$$

$$ \alpha_{i,1,1} = \text{Probability of remaining traveling when traveling}$$
$$\alpha_{i,2,1} = \text{Probability of switching from feeding to traveling}$$

$$\begin{matrix}
  \alpha_{i,1,1} & 1-\alpha_{i,1,1} \\
  \alpha_{i,2,1} & 1-\alpha_{i,2,1} \\
\end{matrix}
$$

##Environment

Behavioral states are a function of local environmental conditions. The first environmental condition is ocean depth. I then build a function for preferential foraging in shallow waters.

It generally follows the form, conditional on behavior at t -1:

$$Behavior_t \sim Multinomial([\phi_{traveling},\phi_{foraging}])$$
$$logit(\phi_{traveling}) = \alpha_{Behavior_{t-1}} + \beta_1 * Ocean_{y[t,]}$$
$$logit(\phi_{foraging}) = \alpha_{Behavior_{t-1}} + \beta_2 * Ocean_{y[t,]}$$

```{r}
traj<-function(gamma=gamma,theta=theta,habitat=habitat,a1_mu,a1_sd,beta_mu,beta_sd,total_time=total_time,step_length){
  
  ##Constants
  #time interval in days, divided into 4 hours intervals
  steps<-length(seq(0,total_time*24,by=step_length))
  
  #generate observations along time interval
  timestep<-sort(runif(300,0,total_time) * 24)  
  
  ##Argos Observation error if 1 degree is 112 km, 0.01 would be ~ 1km error.
  epsilon=0.01
  
  #Position Vector
  xy<-matrix(nrow=steps,ncol=2)

  #Displacement Vector
  d<-matrix(nrow=steps,ncol=2)
  
  #Behavioral States (at time t)
  state<-c()
  
  #Environment
  ocean<-c()

  #indivdual variation intercept
  a1<-c()
  for(x in 1:2){
    a1[x]<-rnorm(1,a1_mu[x],a1_sd[x])  
  }
  
  #indivdual variation slope
  beta<-c()
  for(x in 1:2){
    beta[x]<-rnorm(1,beta_mu[x],beta_sd[x])  
  }
  
  #Probability of staying in behavior
  phi<-matrix(nrow=steps,ncol=2)

  #Mean turning angle
  theta=theta
  
  #Degree of autocorrelation
  gamma=gamma
  
  #Process variance in latitude
  sigma_lat=0.2
  
  #Process variance in longitude
  sigma_lon=0.2
  
  #Correlation in process variance
  rho=0
  
  #Multivariate Normal Variance in random walk
  Sigma<-matrix(nrow=2,ncol=2)
  Sigma[1,1] <- sigma_lon^2
  Sigma[1,2] <- rho * sigma_lon * sigma_lat
  Sigma[2,1] <- rho * sigma_lon * sigma_lat
  Sigma[2,2] <- sigma_lat^2
  
  #Transition matrix for correlated movement
  T<-array(dim=c(steps,2,2))
  
  ##Process Model
  
  ##Initial position##
  xy[1,]<-mvrnorm(mu=c(0,0),Sigma=Sigma)
  
  #First env
  ocean[1]<-extract(habitat,SpatialPoints(data.frame(t(xy[1,]))))
  
  #First behavioral state is a random draw from two probabilities
  lambda<-c()
  lambda[1]<-0.5
  lambda[2]<-1-lambda[1]
  state[1] <- sample(c(1,2),size=1,prob=lambda)
  
  #First step is random walk
  xy[2,]<-mvrnorm(mu=xy[1,],Sigma = Sigma)

  #Draw random walk locations
  for (x in 2:(steps-1)){
    
    #Extract env
    ocean[x]<-extract(habitat,SpatialPoints(data.frame(t(xy[x,]))))
    
    #Behavior
    phi[x,1] <- inv.logit(a1[state[x-1]]+beta[state[x-1]]*ocean[x])
    phi[x,2]<- 1 - phi[x,1]
    state[x] <- sample(c(1,2),size=1,prob=phi[x,])
    
    #Movement correlation matrix
    T[x,1,1] <- cos(theta[state[x]])
    T[x,1,2] <- -sin(theta[state[x]])
    T[x,2,1] <- sin(theta[state[x]])
    T[x,2,2] <- cos(theta[state[x]])
    
    # Add Correlated Displacement
    d[x,]<-xy[x,] + gamma[state[x]] * T[x,,] %*% (xy[x,]  - xy[x-1,])
    
    #next position
    ## Random walk
    xy[x+1,]<-mvrnorm(n=1,mu=d[x,],Sigma=Sigma)
  }
  
  #Format to data frame
  dxy<-data.frame(xy)
  colnames(dxy)<-c("x","y")
  dxy$Step<-1:nrow(dxy)
  
  #Behavior
  dxy$State<-as.factor(c(state,NA))
  
  #Ocean
  #grab last point
  ocean[steps]<-extract(habitat,SpatialPoints(data.frame(t(xy[steps,]))))

  dxy$ocean<-ocean
  levels(dxy$State)<-c("Traveling","Feeding")
  
  #add time label
  dxy$Hour<-seq(0,total_time*24,by=4)
  
  ##Measurement model
  argosdf<-list()
    
  #first location has no interpolation error
  argos_x<-rnorm(1,dxy[dxy$Step==1,c("x")],epsilon)
  argos_y<-rnorm(1,dxy[dxy$Step==1,c("y")],epsilon)
  argosdf[[1]]<-data.frame(Step=1,time=0.5,argos_x,argos_y)

  for (x in 2:steps){
    
      #which time slices sit in interval
      argos_time<-timestep[step_length * (x-1) < timestep & timestep < step_length * x]
      
      #step locations
      trueloc<-dxy[dxy$Step==x,c("x","y")]
      pastloc<-dxy[dxy$Step==(x-1),c("x","y")]

      #for each of those observations
      stepdf<-list()
      for(i in 1:length(argos_time)){
        
        #interpolation distance
        j<-(argos_time[i]-step_length*(x-1))/step_length
        
        #observed locations
        #Add argos noise.
        argos_x<-rnorm(1,(1-j)*pastloc$x+j*trueloc$x,epsilon)
        argos_y<-rnorm(1,(1-j)*pastloc$y+j*trueloc$y,epsilon)
        
        #order dataframe, if there are no observations, fill with NA
        if(!length(argos_time)==0){
          stepdf[[i]]<-data.frame(Step=x,time=argos_time[i],argos_x,argos_y)
        } else{
          stepdf[[i+1]]<-data.frame(Step=x,time=NA,argos_x,argos_y)
        }
      }
      argosdf[[x]]<-rbind_all(stepdf)
    }
  #bind df together
  argosdf<-rbind_all(argosdf)
  
  #merge with true positions
  dxy<-merge(dxy,argosdf,by="Step")

  return(dxy)
}
```

#Simulation

Values come from Jonsen (2005) and Jonsen (2016) fit for foraging seals.

##Behavioral States
### Traveling
$$\gamma_1 = 0.9 = \text{Strong Movement Persistence}$$
$$\theta_1 = 0 = \text{No preference in turns}$$

### Foraging
$$\gamma_2 = 0.1 = \text{Weak Movement Persistence}$$
$$\theta_2 = pi = \text{Many reversals in turns}$$

Essentially, whales travel long straight distances to find food sources, but then stay in those patches for a long time. 

## Environment
Whales tend to travel in deep habitats, slightly weaker effect of ocean depth. The importance of this effect varies by individual.

$$\alpha_{i,1,1} \sim Normal(-2,0.2)$$
$$\beta_{i,1,1} \sim Normal(1,0.1)$$

$$\alpha_{i,2,1} \sim Normal(-2,0.2)$$
$$\beta_{i,2,1} \sim Normal(1,0.1)$$

## Observation Model

Discrete time steps ($x_i$) are related to irregularly spaced observed locations ($X_i$)

$$ x_{n,t,i} = (1-j_{n,t,i})*X_{n,t-1} + j_{n,t,i}*X_{n,t} + \epsilon_{n,t,i}$$

where 

$$ j = \text{The proportion of time interval between locations}$$
$$ \epsilon ~ N(0,\sigma_2)$$

```{r,fig.height=3,fig.width=3.5}
dplot<-function(a1,beta,x){
  y<-inv.logit(a1[1]+beta[1]*x)
  d11<-data.frame(x,y,State="Traveling",Begin="Traveling")
  
  y<-1-inv.logit(a1[1]+beta[1]*x)
  d12<-data.frame(x,y,State="Feeding",Begin="Traveling")

  y<-inv.logit(a1[2]+beta[2]*x)
  d21<-data.frame(x,y,State="Traveling",Begin="Feeding")
  
  y<-1-inv.logit(a1[2]+beta[2]*x)
  d22<-data.frame(x,y,State="Feeding",Begin="Feeding")
  
  d<-rbind_all(list(d11,d12,d21,d22))
}
```

```{r}
#Simulate Data
#Move persistance
gamma=c(0.9,0.1)

#Turning angle
theta=c(0,pi)

#Transitions
#mean
a1_mu=c(-2,-2)

#sd
a1_sd=c(0.2,0.2)

#Effect of ocean depth
#mean
beta_mu=c(1,1)

#sd
beta_sd=c(0.1,0.1)

if(newModel){
  #Simulate environment
  #Spatial auto correlated?
  #m<-raster(mountains(7,0.2))
  #r<-(m-max(m[]))*-3
  #ras2<-raster(xmn=-100,xmx=100,ymn=-100,ymx=100,resolution=0.25,crs=NA)
  #r2<-setExtent(r,ras2)
  #habitat<-resample(r2,ras2)
  
  habitat<-raster(xmn=-100,xmx=100,ymn=-100,ymx=100,resolution=0.25,crs=NA,vals=runif(640000,0,6))
}

#Visualize
d<-dplot(x=seq(min(habitat[]),max(habitat[]),0.1),a1=a1_mu,beta=beta_mu)

ggplot(d,aes(x=x,y=y,col=State)) + geom_line() + geom_point() +  labs(x="Ocean Depth (km)",y="Probability of State") + theme_bw() + facet_wrap(~Begin) + ggtitle("Mean Environmental Relationship")

cr<-colorRampPalette(c("white","black"))
dev.off()
plot(habitat,main='Ocean Depth (km)',col=cr(10))
```

### Multiple trajectories.
```{r}
mxy<-list()

#individuals
ind<-4

#Total time length for each individual track
step.ind<-abs(rnorm(ind,20))

#time for each step
step_length=4

if(newModel){
    for(x in 1:ind){
      mxy[[x]]<-traj(gamma=gamma,theta=theta,habitat=habitat,a1_mu=a1_mu,a1_sd=a1_sd,beta_mu=beta_mu,beta_sd=beta_sd,total_time = step.ind[x],step_length=step_length)
  } 
  mxy<-melt(mxy,id.vars=colnames(mxy[[1]]))
  #Label Individual
  colnames(mxy)[colnames(mxy) %in% 'L1']<-"Animal"
}

#bounding box
bb<-bbox(cbind(mxy$x,mxy$y))

#Habitat dataframe for modeling
crop_habitat<-crop(habitat,bb*2)
habitatdf<-as.data.frame(crop_habitat,xy=TRUE)

ggplot(mxy,aes(x=x,y=y)) + theme_bw() + coord_cartesian(xlim=bb[1,],ylim=bb[2,]) +  geom_raster(data=habitatdf,aes(x=x,y=y,fill=layer)) + scale_fill_continuous("Ocean Depth",low='grey90',high='black')    + geom_path(aes(col=State,group=Animal),size=1) + facet_wrap(~Animal,scales='free')

ggplot(mxy,aes(x=ocean,y=as.numeric(State)-1)) + geom_point() + stat_smooth(method="glm",method.args = list(family="binomial"),alpha=0.1) + scale_y_continuous(breaks=c(0,0.5,1),labels=c("Traveling","0.5","Feeding")) + ylab("Probability") + xlab("Ocean Depth") + theme_bw() + stat_bin_2d() + scale_fill_continuous(low='blue',high='red')
```

#Model Fitting

The goal of the model is to capture the true parameter we simulated above. As we increase complexity, we will be able to monitor the validity of our approach.

```{r}
#Format matrices for jags

#Get the total time, divide into steps, round up
steps_all<-mxy %>% group_by(Animal) %>% summarize(mt=max(time,na.rm=T)) %>% mutate(n=floor(mt/step_length))

#How many observations in each step length segment
mxy$bin<-as.numeric(cut(mxy$time,seq(0,max(steps_all$mt),by=step_length)),include.lowest=TRUE)

# give each step a label
stepmxy<-split(mxy,list(mxy$Step,mxy$Animal),drop = TRUE)
stepmxy<-lapply(stepmxy,function(x){
  x$jStep<-1:nrow(x)
  return(x)
})

mxy<-rbind_all(stepmxy)

#how many observations per individual in each bin
mxy$bin<-factor(mxy$bin,levels=1:max(steps_all$n))
idx<-as.matrix(table(mxy$Animal,mxy$bin))

#make ocean a matrix -> MEAN VALUE -> will this yield a jags error on empty cells?
oc<-acast(mxy,Animal~Step,value.var="ocean",fun.aggregate = mean)

#Individuals
ind=max(mxy$Animal)

#obs array
obs<-melt(mxy,measure.vars=c("argos_x","argos_y"))
obs<-acast(obs,Animal~Step~jStep~variable)

#time diff array
mxy$j<-(mxy$time-step_length*(mxy$Step-1))/step_length
j<-acast(mxy,Animal~Step~jStep,value.var="j")

```

```{r,eval=T}
#source model
source("Bayesian/MultiSpecies.R")

#print model
print.noquote(readLines("Bayesian/MultiSpecies.R"))

#prior cov shape
R <- diag(c(1,1))
argos_prec<-c(0.1,0.1)
data=list(argos=obs,steps=steps_all$n,R=R,ocean=oc,ind=ind,j=j,idx=idx,argos_prec=argos_prec)

#paramters to track
pt<-c("theta","gamma","phi","lalpha_mu","lalpha_tau","lbeta_mu","lbeta_tau","state")

if(newModel){
  system.time(jagM<-jags.parallel(model.file = "Bayesian/Multi_RW.jags",data=data,n.chains=2,parameters.to.save=pt,n.iter=2000,n.burnin=1500,n.thin=2,DIC=FALSE))
}

```

##Chains
```{r,fig.height=10}
#bind chains
pc<-melt(jagM$BUGSoutput$sims.array)
colnames(pc)<-c("Draw","chain","par","value")

#extract parameter name
pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]

#Extract index
splitpc<-split(pc,pc$parameter)

#single index
splitpc[c("lalpha_mu","lalpha_tau","lbeta_mu","lbeta_tau","gamma","theta")]<-lapply(splitpc[c("lalpha_mu","lalpha_tau","lbeta_mu","lbeta_tau","gamma","theta")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,Behavior=sv)
    return(pc)
})

#Double index
splitpc[c("state")]<-lapply(splitpc[c("state")],function(x){
    #As matrices
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+)]"))[,3:4]
    colnames(sv)<-c("Animal","step")
    pc<-data.frame(x,sv)
})

#Three index
splitpc[c("phi")]<-lapply(splitpc[c("phi")],function(x){
#As matrices
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+)]"))[,3:5]
    colnames(sv)<-c("Animal","step","Behavior")
    pc<-data.frame(x,sv)
})

#bind all matrices back together
pc<-rbind_all(splitpc)
rm(splitpc)

#plot all but phi
ggplot(pc[!pc$parameter %in% c("phi","state"),],aes(x=Draw,y=value)) + facet_wrap(~par,scales="free",ncol=2) + geom_line(aes(col=as.factor(chain))) + labs(col="Chain")

```

```{r}
#true values
tv<-melt(list('lalpha_mu[1]'=a1_mu[1],'lalpha_mu[2]'=a1_mu[2],'theta[2]'=theta[2],'theta[1]'=theta[1],'gamma[1]'=gamma[1],'gamma[2]'=gamma[2],'lbeta_mu[1]'=beta_mu[1],'lbeta_mu[2]'=beta_mu[2],'lalpha_tau[1]'=a1_sd[1],'lalpha_tau[2]'=a1_sd[2],'lbeta_tau[2]'=beta_sd[1],'lbeta_tau[2]'=beta_sd[2]))
colnames(tv)<-c("value","par")

#Plot
ggplot(data=pc[!pc$parameter %in% c("phi","state"),],aes(x=value)) + geom_histogram() + facet_wrap(~par,scales='free',ncol=4) + theme_bw() + geom_vline(data=tv,aes(xintercept=value),col='red',linetype='dashed') + ggtitle("Estimated parameters, true values in red")
```

###Compare to priors

```{r}
#add comparison to priors
todraw<-table(pc$par)[1]
pc$Estimate<-"Posterior"

#add priors
#alpha
a1prior<-data.frame(Draw=1:todraw,par='alpha[1]',value=rnorm(todraw,0,1.67),parameter='lalpha_mu',Behavior=1,Estimate="Prior")

a2prior<-data.frame(Draw=1:todraw,par='alpha[2]',value=rnorm(todraw,0,1.67),parameter='lalpha_mu',Behavior=2,Estimate="Prior")

beta1prior<-data.frame(Draw=1:todraw,par='beta[1]',value=rnorm(todraw,0,1.67),parameter='lbeta_mu',Behavior=1,Estimate="Prior")

beta2prior<-data.frame(Draw=1:todraw,par='beta[2]',value=rnorm(todraw,0,1.67),parameter='lbeta_mu',Behavior=2,Estimate="Prior")

gamma1prior<-data.frame(Draw=1:todraw,par='gamma[1]',value=rbeta(todraw,5,2),parameter='gamma',Behavior=1,Estimate="Prior")

gamma2prior<-data.frame(Draw=1:todraw,par='gamma[2]',value=rbeta(todraw,2,5),parameter='gamma',Behavior=2,Estimate="Prior")

prs<-rbind_all(list(a1prior,a2prior,beta1prior,beta2prior,gamma1prior,gamma2prior))

prs$Behavior<-as.factor(prs$Behavior)

allp<-rbind_all(list(pc[!pc$parameter %in% c("phi"),],prs))

ggplot(allp,aes(x=Behavior,y=value,fill=Estimate)) + geom_violin() + facet_wrap(~parameter,scale="free")
```

##Prediction - environmental function

```{r}

#get the posterior estimates of the env function
alphap<-cbind(pc %>% filter(par=='lalpha_mu[1]') %>% .$value,pc %>% filter(par=='lalpha_mu[2]') %>% .$value)

betap<-cbind(pc %>% filter(par=='lbeta_mu[1]') %>% .$value,pc %>% filter(par=='lbeta_mu[2]') %>% .$value)
 
postplot<-list()

for(j in 1:nrow(alphap)){
  postplot[[j]]<-data.frame(dplot(x=mxy$ocean,a1=alphap[j,],beta=betap[j,]),Iteration=j)
  }

postplot<-rbind_all(postplot)

dsum<-postplot %>% group_by(x,Begin,State) %>% summarize(y_mean=mean(y),lower=quantile(y,0.05),upper=quantile(y,0.95))

#overlay true value
trued<-dplot(x=seq(min(habitat[]),max(habitat[]),0.1),a1=a1_mu,beta=beta_mu)

#mean env estimate
ggplot(dsum[,]) + geom_ribbon(aes(x=x,y=y_mean,ymin=upper,ymax=lower,fill=State),alpha=0.5) + theme_bw() + labs(col="Transition",x="Ocean Depth (km)",y="Probability",fill="Transition") + facet_wrap(~Begin) + geom_line(data=trued,aes(x=x,y=y,col=State),linetype='dashed')

```

#Behavioral Prediction

```{r}
#We just need Feeding behavior
behav_chains<-pc[pc$Behavior == 2 & pc$parameter %in% "phi",]

#arrange by time - capitalize
behav_chains$Step<-as.numeric(as.character(behav_chains$step))
behav_chains<-behav_chains %>% arrange(Step)

#Label Behaviors
behav_chains$Behavior<-as.factor(behav_chains$Behavior)
levels(behav_chains$Behavior)<-c("Traveling","Feeding")

#average phi
mean_phi<-behav_chains %>% group_by(Animal,Step) %>% summarize(phi=mean(value))

mxy<-merge(mxy,mean_phi,by=c("Step","Animal"))

ggplot(data=mxy,aes(x=x, y=y,col=phi))+geom_path(size=1.5) + scale_color_continuous(low='blue',high='red') + labs(col="Probability of Feeding") + facet_wrap(~Animal,scales="free") + theme_bw()

#phi and ocean plot
ggplot(data=mxy,aes(x=ocean,y=logit(phi))) + geom_point() + labs(x="Ocean Depth",y="Estimated Log Odds of Feeding") + facet_wrap(~Animal)
```

##Autocorrelation in behavior

```{r}
#create proper time stamp
ggplot(data=mxy,aes(x=time,y=phi,col=ocean)) + geom_line(aes(group=Animal),size=2.5) + labs("Probability of Feeding") + theme_bw() + scale_color_continuous(low='light blue',high='black') + labs(x="Time (hour)",y="Liklihood of Feeding") + facet_wrap(~Animal)
```

##Phase prediction
```{r}
ggplot(mxy[!is.na(mxy$State),],aes(x=phi,fill=State)) + geom_density(alpha=0.8) + labs("Probability of Feeding") + theme_bw() + labs(fill="True State",x="Estimate Probability of Feeding")

#order by time.
mxy<-mxy %>% arrange(Animal,`time`)

#calculate runs
sxy<-split(mxy,mxy$Animal)

state_shade<-lapply(sxy,function(x){
  
  #run lengths
  state_change<-rle(as.numeric(x$State))
  
  #get time for each length switch
  xt<-x$time[cumsum(state_change$lengths[-length(state_change$lengths)])]
  xmax<-x$time[cumsum(state_change$lengths)]
  
  #time limits
  state_shade<-data.frame(xmin=c(0,xt),xmax=xmax,Behavior=state_change$values)
})
names(state_shade)<-1:length(state_shade)
state_shade<-melt(state_shade,c("xmin","xmax","Behavior"))
colnames(state_shade)<-c("xmin","xmax","Behavior","Animal")

state_shade$Behavior<-as.factor(state_shade$Behavior)
levels(state_shade$Behavior)<-c("Traveling","Feeding")
ggplot() + geom_rect(data=state_shade,aes(xmin=xmin,xmax=xmax,fill=Behavior), ymin=-Inf,ymax=Inf) + facet_wrap(~Animal,ncol=1) + theme_bw() + geom_line(data=mxy,aes(x=time,y=phi)) + ylim(0,1) + labs(fill="True Behavior",y="Estimated Probability of Feeding")
```

##Behavioral description

###Average time in phase

* For each draw, create a behavioral sequence, calculate the average runs for each behavior

## Predicted Run Length
```{r}
statec<-pc %>% filter(parameter=='state')

#split by animal and draw and chain
astate<-split(statec,list(statec$Animal,statec$Draw,statec$chain))

#get run length

runs<-lapply(astate,function(x){
  #arrange by step
  #calculate run length
  state_change<-rle(as.numeric(x$value))
  runl<-data.frame(Animal=unique(x$Animal),runs=state_change$lengths,Behavior=state_change$values)
  runl$Behavior<-as.factor(runl$Behavior)
  levels(runl$Behavior)<-c("Traveling","Feeding")
  
  return(runl)
})

runs<-rbind_all(runs)

ggplot(runs,aes(x=as.numeric(runs),fill=Behavior)) + geom_histogram(position='dodge') + scale_x_continuous(breaks=seq(0,max(runs$runs),by=10)) + labs(x="Run Length") + ggtitle("Predicted Behavior Duration") + theme_bw()
```

```{r,fig.height=3,fig.with=3.5}
mr<-runs %>% group_by(Behavior) %>% summarize(mean=mean(runs),lower=quantile(runs,0.05),upper=quantile(runs,0.95))
ggplot(mr,aes(x=Behavior,y=mean,ymin=lower,ymax=upper)) + geom_pointrange(size=2) + theme_bw() + ggtitle("Average Duration of Behavior") + labs(y="Steps")
```

###Proportional Hazards

Survival analysis typically examines the relationship between time to death as a function of covariates. From this we can get the instantaneous rate of death at time t f(t), which is the cumulative distribution of the likelihood of death.

Let T represent survival time.

$$ P(t) = Pr(T<t)$$ 
with a pdf
$$p(t) = \frac{dP(t)}{dt}$$

The instantaneous risk of death at time t (h(t)), conditional on survival to that time:

$$ h(t) = \lim{\Delta_t\to 0} \frac{Pr[(t<T<t + \Delta_t)]|T>t}{\delta t}$$

with covariates:
$$log (h_i(t)) = \alpha + \beta_i *x$$

The cox model has no intercept, making it semi-parametric
$$ log(h_i(t)) = h_0(t) + \beta_1 * x$$

```{r}
#Create a 'bout' encoding for every time it changes
statec[cumsum(runs$runs),"Bout"]<-seq(1,nrow(runs),1)

#Fill missing NA values
repeat.before = function(x) {   # repeats the last non NA value. Keeps leading NA
    ind = which(!is.na(x))      # get positions of nonmissing values
    if(is.na(x[1]))             # if it begins with a missing, add the 
          ind = c(1,ind)        # first position to the indices
    rep(x[ind], times = diff(   # repeat the values at these indices
       c(ind, length(x) + 1) )) # diffing the indices + length yields how often 
}                               

statec$Bout<-repeat.before(statec$Bout)

feedr<-runs[runs$Behavior %in% "Feeding",]

#Every run eventually ends in a switch (that's what makes it a run)
feedr$status<-1

#survival analysis
tf<-coxph(Surv(time=feedr$runs,event=feedr$status)~feedr$Animal)
summary(tf)
plot(survfit(tf),ylab="Proportion still feeding",xlab="Duration (steps)")
```


```{r}
save.image("MultiSpecies.RData")
```
