---
title: "Antarctic Whale Project: Patch Dynamics"
author: "Ben Weinstein"
date: "`r Sys.time()`"
output: 
  html_document:
    keep_md: true
    number_sections: true
    toc: true
    theme: spacelab
---

#Aim


```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width = 6,fig.height = 4,fig.align='center')
library(boot)
library(raster)
library(reshape2)
library(ggplot2)
library(survival)
library(MASS)
library(ggmap)
library(dplyr)
```

```{r,eval=T}
#Load in data
feedr<-read.csv("OutData/Runs.csv")
```


#Proportional Hazards

Survival analysis typically examines the relationship between time to death as a function of covariates. From this we can get the instantaneous rate of death at time t f(t), which is the cumulative distribution of the likelihood of death.

Let T represent survival time.

$$ P(t) = Pr(T<t)$$ 
with a pdf
$$p(t) = \frac{dP(t)}{dt}$$

The instantaneous risk of death at time t (h(t)), conditional on survival to that time:

$$ h(t) = \lim{\Delta_t\to 0} \frac{Pr[(t<T<t + \Delta_t)]|T>t}{\Delta t}$$

with covariates:
$$log (h_i(t)) = \alpha + \beta_i *x$$

The cox model has no intercept, making it semi-parametric
$$ log(h_i(t)) = h_0(t) + \beta_1 * x$$

```{r}
#run functions
runf<-function(x){
  #arrange by step
  #calculate run length
  state_change<-rle(as.numeric(x))
  runl<-data.frame(runs=state_change$lengths,Behavior=state_change$values)
  runl$Behavior<-as.factor(runl$Behavior)
  levels(runl$Behavior)<-c("Traveling","Feeding")
  
  return(runl)
}

endtrack<-function(x){
  x[nrow(x),"status"]<-0
  return(x)
}
```

#Random foraging.

If behavioral states were distributed at random, what would the survival and hazard functions look like?


```{r}
##Create random runs

#total number of hours
th<-sum(feedr$hours)
#
random_draw<-rbinom(th,1,0.5)
random_run<-runf(random_draw)
random_run$status<-1

#last track is censored
random_run<-endtrack(random_run)
```


```{r}
##Fit Cox Regression

#survival analysis for just feeding
rf<-coxph(Surv(time=random_run$runs,event=random_run$status,type="right")~1)
rfit<-survfit(rf)
```


```{r}
##Survival Curve

#plot
splot1<-data.frame(x=rfit$time,surv=rfit$surv,lower=rfit$lower,upper=rfit$upper)

ggplot(splot1,aes(x=x,y=surv,ymin=lower,ymax=upper)) + geom_step() + geom_ribbon(alpha=0.3) + theme_bw() + labs(x="Time",y="Proportion still feeding") + ggtitle("Random Foraging")
```

## Hazard function

$$\text{Cumulative Hazard} = -log(Survival)$$

```{r}
hplot1<-data.frame(x=rfit$time,chaz=-log(rfit$surv),lower=-log(rfit$lower),upper=-log(rfit$upper))

ggplot(hplot1,aes(x=x,y=chaz,ymin=lower,ymax=upper)) + geom_step() + geom_ribbon(alpha=0.3) + theme_bw() + labs(x="Time",y="Cumulative Hazard")
```

# Observed Data


```{r}
#survival analysis for just feeding
tf<-coxph(Surv(time=feedr$hours,event=feedr$status,type="right")~1)
fit<-survfit(tf)
```


## Survival

```{r}
#plot
splot2<-data.frame(x=fit$time,surv=fit$surv,lower=fit$lower,upper=fit$upper)
ggplot(splot2,aes(x=x,y=surv,ymin=lower,ymax=upper)) + geom_step() + geom_ribbon(alpha=0.3) + theme_bw() + labs(x="Time",y="Proportion still feeding")
```

## Cumulative Hazard

```{r}
#plot
hplot2<-data.frame(x=fit$time,chaz=-log(fit$surv),lower=-log(fit$lower),upper=-log(fit$upper))
ggplot(hplot2,aes(x=x,y=chaz,ymin=lower,ymax=upper)) + geom_step() + geom_ribbon(alpha=0.3) + theme_bw() + labs(x="Time",y="Cumulative Hazard")
```

#Comparison

## Survival

```{r}
survival<-list(splot1,splot2)
names(survival)<-c("Random","Observed")
survival<-melt(survival,id.vars=colnames(splot1))
ggplot(survival,aes(x=x,y=surv,ymin=lower,ymax=upper,col=L1,fill=L1)) + geom_step() + geom_ribbon(alpha=0.3) + theme_bw() + labs(x="Time",y="Cumulative Hazard") + labs(col="Data",fill="Data")
```

## Hazards

```{r}
hazards<-list(hplot1,hplot2)
names(hazards)<-c("Random","Observed")
hazards<-melt(hazards,id.vars=colnames(hplot1))
ggplot(hazards,aes(x=x,y=chaz,ymin=lower,ymax=upper,col=L1,fill=L1)) + geom_step() + geom_ribbon(alpha=0.3) + theme_bw() + labs(x="Time",y="Proportion stil feeding") + labs(col="Data",fill="Data")
```

