---
title: "Antarctic Whale Project: Single Species"
author: "Ben Weinstein"
date: "`r Sys.time()`"
output: 
  html_document:
    keep_md: yes
    number_sections: yes
    theme: spacelab
    toc: yes
    pandoc_args: [
    "-M2GB", "+RTS", "-K64m", "-RTS"
    ]
---

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width = 11,fig.height = 5,cache=F)
library(boot)
library(reshape2)
library(ggplot2)
library(MASS)
library(ggmap)
library(dplyr)
library(chron)
library(gridExtra)
library(stringr)
library(R2jags)
library(move)
library(tidyr)
library(maps)
library(maptools)
library(raster)

#New model to be run, flag turned off if just updating.
newModel<-T
```

```{r,eval=F}
#Load in data
load("SingleSpecies.RData")
#jags$recompile()
newModel<-F
```

```{r}
#memory function

# improved list of objects
.ls.objects <- function (pos = 1, pattern, order.by,
                        decreasing=FALSE, head=FALSE, n=5) {
    napply <- function(names, fn) sapply(names, function(x)
                                         fn(get(x, pos = pos)))
    names <- ls(pos = pos, pattern = pattern)
    obj.class <- napply(names, function(x) as.character(class(x))[1])
    obj.mode <- napply(names, mode)
    obj.type <- ifelse(is.na(obj.class), obj.mode, obj.class)
    obj.prettysize <- napply(names, function(x) {
                           capture.output(format(utils::object.size(x), units = "auto")) })
    obj.size <- napply(names, object.size)
    obj.dim <- t(napply(names, function(x)
                        as.numeric(dim(x))[1:2]))
    vec <- is.na(obj.dim)[, 1] & (obj.type != "function")
    obj.dim[vec, 1] <- napply(names, length)[vec]
    out <- data.frame(obj.type, obj.size, obj.prettysize, obj.dim)
    names(out) <- c("Type", "Size", "PrettySize", "Rows", "Columns")
    if (!missing(order.by))
        out <- out[order(out[[order.by]], decreasing=decreasing), ]
    if (head)
        out <- head(out, n)
    out
}

# shorthand
lsos <- function(..., n=10) {
    .ls.objects(..., order.by="Size", decreasing=TRUE, head=TRUE, n=n)
}

#ggplot theme
mytheme<-theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),axis.title.x=element_blank(),axis.title.y=element_blank())
```

```{r}
#read data
mdat<-read.csv("InputData/FilteredData.csv",row.names=1)
#standardize column names to match the simulation
#Create an animal tag.
mxy <- as(mdat, "data.frame")
mxy$Animal<-mxy$individual.local.identifier
mxy$x<-mxy$location.long
mxy$y<-mxy$location.lat

#make sure to remove minke whales, should be already gone, but justfor safety
minke<-c("131117","131118","131120","154184")
mxy<-mxy[!mxy$individual.local.identifier %in% minke,]

#grab set of animals
mxy<-mxy[mxy$Animal %in% c("131143","131142","131144","131147","121210"),]

#empty coordinates
mxy<-mxy[!is.na(mxy$x),]

#crop by extent
d<-SpatialPointsDataFrame(cbind(mxy$x,mxy$y),data=mxy,proj4string=CRS("+proj=longlat +datum=WGS84"))

cropoly<-readShapePoly("InputData/CutPolygon.shp",proj4string=CRS("+proj=longlat +datum=WGS84"))

b<-d[!is.na(d %over% cropoly)[,2],]

mxy<-b@data

#set datestamp
mxy$timestamp<-as.POSIXct(mxy$timestamp)

#month and year columns
mxy$Month<-months(mxy$timestamp)
mxy$Year<-years(mxy$timestamp)

#Only austral sping and summer
mxy<-mxy[mxy$Month %in% month.name[1:7],]

#remove empty timestamps
mxy<-mxy[!is.na(mxy$timestamp),]

#remove duplicates
mxy<-mxy[!duplicated(data.frame(mxy$timestamp,mxy$Animal)),]
mxy<-mxy[!mxy$ETOPO1.Elevation>0,]


```

```{r}
#hold copy of orig
oxy<-mxy

#remake spatial object
d<-SpatialPointsDataFrame(cbind(mxy$x,mxy$y),data=mxy,proj4string=CRS("+proj=longlat +datum=WGS84"))

#Plot
pp<-c(mean(mxy$x),mean(mxy$y))
m <- get_map(location=pp,source="google",zoom=4,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal)),size=.5) + scale_color_discrete(guide="none") 
```

##By Month

```{r,fig.height=12}
mxy$Month<-factor(mxy$Month,levels=month.name)
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal)),size=.5)  + facet_wrap(~Month,nrow = 2) + scale_color_discrete(guide="none") + labs(x="",y="") + mytheme
ggsave("Figures/MonthTracks.jpeg",height=6, width=9)
```

##CCAMLR Units
```{r}
ccamlr<-readShapePoly("InputData/SSMU_2013/CROPCCAMLR.shp")

ccamlr$Unit<-gsub(x=ccamlr$SSMUname,pattern="Antarctic Peninsula ",rep="")

fccamlr<-fortify(ccamlr,region="Unit")

m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color = "bw")

ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal)),size=.5) + scale_color_discrete(guide="none") + geom_polygon(data=fccamlr,aes(x=long,y=lat,group=id),col="black",alpha=0.01) + mytheme
```

##Distance

```{r,fig.height=3,eval=F}
moved<-move(x=mxy$x,y=mxy$y,time=mxy$timestamp,proj=CRS("+proj=longlat +ellps=WGS84"),animal=mxy$Animal,data=as.data.frame(mxy),sensor=mxy$Animal,removeDuplicatedTimestamps=T)
dstat<-melt(move::distanceSummary(moved))

dstat<-droplevels(dstat[dstat$variable %in% c("TravDist","AverDist","FarthDist"),])
levels(dstat$variable)<-c("Total Distance","Average Distance Between Points","Farthest Distance Between Points")
ggplot(dstat,aes(value/1000)) + geom_histogram() + facet_wrap(~variable,scales="free") + labs(x="Kilometers")
```

##Time 
```{r,fig.height=3,eval=F}
tstat<-move::timeSummary(moved)
tstat<-melt(tstat,measure.vars=colnames(tstat[[1]]))
tstat<-droplevels(tstat[tstat$variable %in% c("Duration","AverDur"),])
levels(tstat$variable)<-c("Total Time (Hours)","Average Time Between Points (Hours)")
ggplot(tstat,aes(value)) + geom_histogram() + facet_wrap(~variable,scales="free")
```

##Velocity
```{r,fig.height=3,eval=F}
vstat<-move::speedSummary(moved)
vstat<-melt(vstat,measure.vars=colnames(vstat[[1]]))
levels(vstat$variable)<-c("Average Speed (m/s)","Variance in Speed (m/s)","Max Speed (m/s)")
ggplot(vstat,aes(value)) + geom_histogram() + facet_wrap(~variable,scales="free")
```

##Angles

```{r,fig.height=2,eval=F}
astat<-move::angleSummary(moved)
astat<-melt(astat,measure.vars=colnames(astat[[1]]))
astat<-droplevels(astat[astat$variable %in% "AverAzimuth",])
levels(astat$variable)<-"Turning Angle (degrees)"
ggplot(astat,aes(value)) + geom_histogram() + facet_wrap(~variable,scales="free")

#remove move object
rm(moved)
```

#Correlated random walk

*Process Model*

$$ d_{t} \sim T*d_{t-1} + Normal(0,\Sigma)$$
$$ x_t = x_{t-1} + d_{t} $$

## Parameters

For each individual:

$$\theta = \text{Mean turning angle}$$
$$\gamma = \text{Move persistence} $$

For both behaviors process variance is:
$$ \sigma_{latitude} = 0.1$$
$$ \sigma_{longitude} = 0.1$$

##Behavioral States

$$ \text{For each individual i}$$
$$ Behavior_1 = \text{traveling}$$
$$ Behavior_2 = \text{foraging}$$

$$ \alpha_{i,1,1} = \text{Probability of remaining traveling when traveling}$$
$$\alpha_{i,2,1} = \text{Probability of switching from Foraging to traveling}$$

$$\begin{matrix}
  \alpha_{i,1,1} & 1-\alpha_{i,1,1} \\
  \alpha_{i,2,1} & 1-\alpha_{i,2,1} \\
\end{matrix}$$


With the probability of switching states:

$$logit(\phi_{traveling}) = \alpha_{Behavior_{t-1}}$$

$$\phi_{foraging} = 1 - \phi_{traveling} $$

##Continious tracks

The transmitter will often go dark for 10 to 12 hours, due to weather, right in the middle of an otherwise good track. The model requires regular intervals to estimate the turning angles and temporal autocorrelation. As a track hits one of these walls, call it the end of a track, and begin a new track once the weather improves. We can remove any micro-tracks that are less than three days.
Specify a duration, calculate the number of tracks and the number of removed points. Iteratively.

```{r}
##Time is the beginning of the first point.
step_length=12

sxy<-split(mxy,mxy$Animal)

#time diff function
timed<-function(d,step_length){
  d$j[1]<-0
  for (x in 2:nrow(d)){
    d$j[x]<-as.numeric(difftime(as.POSIXct(d$timestamp[x]),as.POSIXct(d$timestamp[x-1]),units="mins"))/(step_length*60)
  }
  
  #Split out track endings
  ends<-c(1,which(d$j>1),nrow(d))

  for(w in 2:length(ends)){
    d[ends[w-1]:ends[w],"Track"]<-w-1
  }
  
  #remove tracks that are shorter than three days
  track_time<-d %>% group_by(Track) %>% summarize(mt=difftime(max(as.POSIXct(timestamp)),min(as.POSIXct(timestamp)),units="hours")) %>% filter(mt>=24) %>% .$Track
  
  d<-d[d$Track %in% track_time,]
  
  #renumber the tracks
  d$Track<-as.numeric(as.factor(d$Track))
  return(d)
  }

sxy<-lapply(sxy,timed,step_length=step_length)

#Format matrices for jags
mxy<-rbind_all(sxy)

######recode whales
#mxy$Animal<-as.numeric(as.factor(mxy$Animal))

sxy<-split(mxy,list(mxy$Animal,mxy$Track),drop=TRUE)

sxy<-lapply(sxy,function(x){
#How many observations in each step length segment
x$step<-as.numeric(cut(as.POSIXct(x$timestamp),"12 hours"))
return(x)
})

mxy<-rbind_all(sxy)

#refactor animal
mxy$Animal<-as.numeric(as.factor(mxy$Animal))
```

```{r}
#total number of steps per track/animal
steps_all<-mxy %>% group_by(Animal,Track) %>% summarize(n=length(unique(step)))

# give each step a label
mxy<-mxy %>% group_by(Animal,Track,step) %>% mutate(jStep=1:n())

#Cast time array
j<-acast(mxy,Animal~Track~step~jStep,value.var="j")

#how many observations per individual in each step
mxy$step<-factor(mxy$step,levels=1:max(steps_all$n))
idx<-melt(table(mxy$Animal,mxy$Track,mxy$step))
colnames(idx)<-c("Animal","Track","step","jStep")
idx<-acast(data=idx,Animal~Track~step)

#month array
mxy$MonthF<-as.numeric(factor(mxy$Month,levels=month.name))

MonthA<-acast(mxy,Animal~Track~step,value.var="MonthF",fun.aggregate = min)
MonthA[!is.finite(MonthA)]<-NA

#Individuals
ind=length(unique(mxy$Animal))

#tracks per indivudal
tracks<-mxy %>% group_by(Animal) %>% summarize(tracks=length(unique(Track))) %>% .$tracks

#steps per track
steps<-acast(steps_all,Animal~Track,value.var="n")

#obs array
obs<-melt(mxy,measure.vars=c("x","y"))
obs<-acast(obs,Animal~Track~step~jStep~variable)
```

How did the filter change the extent of tracks?

```{r}
#Plot
mxy<-mxy %>% arrange(Animal,Track,step,jStep)

pp<-c(mean(mxy$x),mean(mxy$y))
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal),group=paste(Animal,Track)),size=.5) + scale_color_discrete(guide="none") + mytheme
```

```{r,fig.height=15}
ggplot() +  geom_path(data=oxy,col='blue',aes(x=x,y=y)) +  geom_path(data=mxy,col='red',aes(x=x,y=y),size=2,alpha=0.7)  + facet_wrap(~individual.local.identifier,scales="free",ncol=4)
```

Look at the observations were defined into tracks.

```{r,fig.height=20}
cl<-rainbow(100)
cl<-sample(cl)
p<-ggplot() + geom_path(data=mxy, aes(x=x, y=y,group=1),size=0.1,col="black") + labs(col="step")
p<-p+geom_path(data=mxy, aes(x=x, y=y,col=as.numeric(step),group=paste(Track,step)),size=1) + labs(x="",y="") + theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),axis.ticks.x=element_blank()) + scale_color_gradientn(colours = cl) + theme_bw() + facet_wrap(~Animal,scales="free") 
p + mytheme
```

```{r,fig.height=3,fig.width=5}
ggplot(data=steps_all,aes(x=n*step_length/24)) + geom_histogram() + labs(x="Days") + ggtitle("Track Length")
ggplot(data=steps_all,aes(x=Track)) + geom_histogram() + labs(x="Subtracks per Animal")
```

```{r,child="Bayesian/MultiSpecies.R",eval=T}
```

```{r,eval=T}
#source jags file
source("Bayesian/MultiSpecies.R")

#prior cov shape
R <- diag(c(1,1))
data=list(argos=obs,steps=steps,R=R,ind=ind,j=j,idx=idx,tracks=tracks,Month=MonthA,Months=max(MonthA,na.rm=T))

#paramters to track
pt<-c("theta","gamma","phi","alpha_mu","state")

if(newModel){
  system.time(jagM<-jags.parallel(model.file = "Bayesian/Multi_RW.jags",data=data,n.chains=2,parameters.to.save=pt,n.iter=1000,n.burnin=900,n.thin=2,DIC=FALSE))
}

```

```{r,eval=F}
#update function
recompile(jagM)
jagM <- update(jagM,n.iter=300,n.burnin=0,n.thin=0) 
```

##Chains
```{r,eval=T}

#check memory size
lsos()

#delete jags objects

#rm(data)
#rm(argos)
#rm(obs)
#rm(j)
gc()

#bind chains
pc<-melt(jagM$BUGSoutput$sims.array)

rm(jagM)

colnames(pc)<-c("Draw","chain","par","value")

#extract parameter name
pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]

#Extract index
splitpc<-split(pc,pc$parameter)

#single index
splitpc[c("alpha_mu","gamma","theta")]<-lapply(splitpc[c("alpha_mu","gamma","theta")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,Behavior=sv)
    return(pc)
})

#double index

#Three index
splitpc[c("phi")]<-lapply(splitpc[c("phi")],function(x){
#As matrices
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+),(\\d+)]"))[,3:6]
    colnames(sv)<-c("Animal","Track","step","Behavior")
    pc<-data.frame(x,sv)
})

#State index
splitpc[c("state")]<-lapply(splitpc[c("state")],function(x){
#As matrices
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+)]"))[,3:5]
    colnames(sv)<-c("Animal","Track","step")
    pc<-data.frame(x,sv)
})

#bind all matrices back together
pc<-rbind_all(splitpc)
rm(splitpc)

```

```{r,fig.height=8,fig.width=14}
#plot all but phi and state
ggplot(pc[!pc$parameter %in% c("phi","state"),],aes(x=Draw,y=value)) + facet_wrap(~par,scales="free",ncol=3) + geom_line(aes(col=as.factor(chain))) + labs(col="Chain")
```

```{r,eval=F}
#write.csv
#write.csv(pc,"OutData/pars.csv")
#pc<-read.csv("OutData/pars.csv")
pc$Behavior<-as.factor(pc$Behavior)
```

```{r,fig.height=10}
#Plot
ggplot(data=pc[!pc$parameter %in% c("phi","state"),],aes(x=value)) + geom_histogram() + facet_wrap(~par,scales='free',ncol=3) + theme_bw()  + ggtitle("Estimated parameters")
```

Look at the convergence of phi, just for an example

```{r}
pc %>% filter(Animal==1,step %in% 2:10,Track==1,Behavior==2) %>% filter(parameter=="phi") %>% ggplot(.,aes(x=Draw,y=value,col=as.factor(chain))) + geom_line() + facet_wrap(~step)
```

Overall relationship between phi and state, nice test of convergence.

```{r}
a<-pc %>% filter(parameter %in% c("state","phi")) %>% filter((parameter == "phi"&Behavior==2)|parameter =="state") %>% dplyr::select(Draw,chain,Animal,Track,step,value,parameter) %>% spread(key=parameter,value=value)
ggplot(a,aes(x=phi,fill=as.factor(state-1))) + geom_density(alpha=0.7) + labs(fill="Behavior")

```

###Compare to priors

```{r}
#add comparison to priors
todraw<-table(pc$par)[1]
pc$Estimate<-"Posterior"

#add priors
#alpha
a1prior<-data.frame(Draw=1:todraw,par='alpha[1]',value=rnorm(todraw,0,1.67),parameter='alpha_mu',Behavior=1,Estimate="Prior")

a2prior<-data.frame(Draw=1:todraw,par='alpha[2]',value=rnorm(todraw,0,1.67),parameter='alpha_mu',Behavior=2,Estimate="Prior")

gamma1prior<-data.frame(Draw=1:todraw,par='gamma[1]',value=rbeta(todraw,5,2),parameter='gamma',Behavior=2,Estimate="Prior")

gamma2prior<-data.frame(Draw=1:todraw,par='gamma[2]',value=rbeta(todraw,5,2) * rbeta(todraw,1,1),parameter='gamma',Behavior=1,Estimate="Prior")

prs<-rbind_all(list(a1prior,a2prior,gamma1prior,gamma2prior))

prs$Behavior<-as.factor(prs$Behavior)

allp<-rbind_all(list(pc[!pc$parameter %in% c("phi","state"),],prs))

ggplot(allp,aes(x=Behavior,y=value,fill=Estimate)) + geom_violin() + facet_wrap(~parameter,scale="free")
rm(allp)
```

## Parameter Summary
```{r,fig.height=10,fig.width=15}
psum<-pc %>% filter(!parameter %in% c("phi","state")) %>% group_by(parameter,par) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

as.data.frame(psum)

ggplot(psum,aes(x=par,y=mean,ymax=upper,ymin=lower)) + geom_linerange() + geom_point() + facet_wrap(~parameter,scales="free",ncol=2)
```

#Behavioral Prediction

```{r}
#mean transition probability
mean_phi<-pc %>% filter(parameter=="phi",Behavior==2) %>% mutate(step=as.numeric(as.character(step))) %>% group_by(Animal,Track,step) %>% summarize(phi=mean(value)) %>% arrange(Animal,Track,step)

#mode state for each position
mode_state<-pc %>% filter(parameter %in% "state") %>% mutate(bstate=ifelse(value==1,"Traveling","Area-restricted Search")) %>% group_by(Animal,Track,step,bstate) %>% summarize(n=n()) %>% spread(key=bstate,value=n,fill=0)%>% mutate(ratio=`Area-restricted Search`/Traveling) %>% mutate(state=as.character(ifelse(ratio>=1,"Area-restricted Search","Traveling")))

mxy<-merge(mxy,mean_phi,by=c("step","Track","Animal"))
mxy<-merge(mxy,mode_state,by=c("step","Track","Animal"))

#ensure order for plotting
mxy<-mxy %>% arrange(Animal,Track,step,jStep)

#make factor
mxy$bstate<-factor(mxy$state)

#make phi a factor
mxy$phistate<-cut(mxy$phi,c(0,0.5,1),labels=c("Traveling","Area-restricted Search"))

```

Relationship between phi and state

```{r}
ggplot(mxy,aes(x=phi,fill=bstate)) + geom_density(alpha=0.5) + labs(fill="Behavior",x="Probability of switching to ARS")
```

##Spatial Prediction

```{r}
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color="bw")

ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=bstate,group=paste(Animal,Track)),size=.2)  + geom_path(data=fccamlr,aes(x=long,y=lat,group=id),col="black",size=1) + mytheme + labs(col="Behavior")

ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=phistate,group=paste(Animal,Track)),size=.2) + geom_path(data=fccamlr,aes(x=long,y=lat,group=id),col="black",size=1)
ggsave("Figures/SpatialPred.jpg",height=5,width=5)

```

## By individual

```{r}
ggplot()+geom_path(data=mxy, aes(x=x, y=y,col=bstate,group=paste(Animal,Track)),size=.5)  + facet_wrap(~individual.local.identifier,scales="free") + mytheme

ggplot()+geom_path(data=mxy, aes(x=x, y=y,col=phistate,group=paste(Animal,Track)),size=.5) + facet_wrap(~individual.local.identifier,scales="free") + mytheme 
```

Overlay phi and state

```{r}
ggplot()+geom_path(data=mxy, aes(x=x, y=y,col=(phistate==bstate),group=paste(Animal,Track)),size=.5) + facet_wrap(~individual.local.identifier,scales="free") + mytheme 
```

## Compared to CMLRR regions

```{r}
ggmap(m)+ geom_path(data=fccamlr,aes(x=long,y=lat,group=id),alpha=0.5,col="black")+geom_path(data=mxy, aes(x=x, y=y,col=phi,group=paste(Animal,Track)),size=.5)   + scale_color_continuous(low='blue',high='red') + labs(col="Probability of Foraging") 
ggsave("Figures/SpatialPred_regions.jpg",height=5,width=5)
```

##Autocorrelation in behavior

```{r,fig.height=20,fig.width=13}
#create proper time stamp
ggplot(data=mxy,aes(x=as.POSIXct(timestamp),y=phistate)) + geom_line(aes(group=Track),size=1) + theme_bw() + scale_color_continuous(low='light blue',high='black') + labs(x="Time",y="Behavior") + facet_wrap(~individual.local.identifier,ncol=3,scales="free")
```

##Location of Behavior

```{r}
msp<-SpatialPointsDataFrame(cbind(mxy$x,mxy$y),data=mxy[,c("x","y","Animal","phi","timestamp")],proj=CRS("+proj=longlat +datum=WGS84"))

r<-raster(msp,ext=extent(c(-75,-50,-70,-57.5)))
res(r)<-0.1

m<-rasterize(x=msp,y=r,field="phi")

#plotting
feedmap <- data.frame(rasterToPoints(m))

temp <- get_map(location=bbox(m),source="google",zoom=5,maptype="satellite",color = "bw",scale = 2)

```

# Overlap with Krill Fishery
```{r}
krill<-read.csv("InputData/CCAMLR_aggregated_catch_C1.csv")
ksp<-SpatialPointsDataFrame(cbind(krill$GridMidpointDegreeLon,krill$GridMidpointHalfDegreeLat),krill)

#get the ccamlr units
kdat<-ksp %over% ccamlr
ksp<-SpatialPointsDataFrame(cbind(krill$GridMidpointDegreeLon,krill$GridMidpointHalfDegreeLat),data.frame(kdat,data.frame(ksp)))
```

#Time spent in grid cell

```{r}
#raster
r<-raster(extent(d)*c(1.1,1.1),nrow=100,ncol=100)

#res(r)<-0.25
#put cellindex number
r[]<-1:ncell(r)

sxy<-split(mxy,mxy$Animal,drop=T)
sxy<-lapply(sxy,function(x) {x$Index<-1:nrow(x); return(x)})

#plot(rasterize(j,r,field="Index",fun='count'))

#time in each cell
timecell<-function(j){
  
  rlej<-rle(x=diff(j$Index))
  position<-c(1,cumsum(rlej$lengths))
  timestack<-c()
  
  #if only one position
  if(length(position)==1){
    timestack[1]<-0
    
  } else {
    
    #time difference between first and last point in the cell
      for(x in 2:length(position)){
        timed<-j[position[x-1]:position[x],]
        timestack[x-1]<-difftime(max(timed$timestamp),min(timed$timestamp),units="days")
    }
  }

  out<-data.frame(Sum=sum(timestack))
  return(out)
}

#loop through all animals
sptracks<-lapply(sxy,function(x){
  k<-SpatialPointsDataFrame(cbind(x$x,x$y),data=x,proj4string=CRS("+proj=longlat +datum=WGS84"))
  
  #get the cells that points are in
  k$cellvalue<-raster::extract(r,k)
  
  #for each cell value, perform time function
  sptrack<-as.data.frame(k) %>% group_by(Animal,cellvalue) %>% do(timecell(.))
  return(sptrack)}
)

sptracks<-rbind_all(sptracks)

#get sum across animals
cellvals<-sptracks %>% group_by(cellvalue) %>% summarize(Sum=sum(Sum))

#is there a better way to loop through?
r[]<-NA
for(x in 1:nrow(cellvals)){
  r[cellvals$cellvalue[x]]<-cellvals$Sum[x]
}
s<-stack(r)
names(s)[[1]]<-"All Behavior"
```

```{r}
#extract
ksp$ResidencyTime<-raster::extract(r,y=ksp)
#write to file
ggplot(as.data.frame(ksp)) + geom_point(aes(x=ResidencyTime,y=log(C1KRIcatchKG)))  + theme_bw() + labs(x="Residency Time (Days)",y="Krill Catch (Log Metric Tons)")
```

```{r}
traveling<-mxy %>% filter(phistate=="Traveling")
foraging<-mxy %>% filter(phistate=="Area-restricted Search")
```

```{r}
#raster
if(!nrow(traveling)==0){
rtraveling<-raster(extent(d)*c(1.1,1.1),nrow=100,ncol=100)

#res(rtraveling)<-0.25
#put cellindex number
rtraveling[]<-1:ncell(rtraveling)


sxy<-split(traveling,traveling$Animal)
sxy<-lapply(sxy,function(x) {x$Index<-1:nrow(x); return(x)})

#plot(rasterize(j,rtraveling,field="Index",fun='count'))

#time in each cell
timecell<-function(j){
  
  rlej<-rle(x=diff(j$Index))
  position<-c(1,cumsum(rlej$lengths))
  timestack<-c()
  
  #if only one position
  if(length(position)==1){
    timestack[1]<-0
    
  } else {
    
    #time difference between first and last point in the cell
      for(x in 2:length(position)){
        timed<-j[position[x-1]:position[x],]
        timestack[x-1]<-difftime(max(timed$timestamp),min(timed$timestamp),units="days")
    }
  }

  out<-data.frame(Sum=sum(timestack))
  return(out)
}

#loop through all animals
sptracks<-lapply(sxy,function(x){
  k<-SpatialPointsDataFrame(cbind(x$x,x$y),data=x,proj4string=CRS("+proj=longlat +datum=WGS84"))
  
  #get the cells that points are in
  k$cellvalue<-raster::extract(rtraveling,k)
  
  #for each cell value, perform time function
  sptrack<-as.data.frame(k) %>% group_by(Animal,cellvalue) %>% do(timecell(.))
  return(sptrack)}
)

sptracks<-rbind_all(sptracks)

#get sum across animals
cellvals<-sptracks %>% group_by(cellvalue) %>% summarize(Sum=sum(Sum))

#is there a better way to loop through?
rtraveling[]<-NA
for(x in 1:nrow(cellvals)){
  rtraveling[cellvals$cellvalue[x]]<-cellvals$Sum[x]
}
s[[2]]<-rtraveling
names(s)[[2]]<-"Traveling"
}
```

## ARS

```{r}
if(!nrow(foraging)==0){
#raster
rforaging<-raster(extent(d)*c(1.1,1.1),nrow=100,ncol=100)

#res(r)<-0.25
#put cellindex number
rforaging[]<-1:ncell(rforaging)

sxy<-split(foraging,foraging$Animal)
sxy<-lapply(sxy,function(x) {x$Index<-1:nrow(x); return(x)})

#plot(rasterize(j,r,field="Index",fun='count'))

#time in each cell
timecell<-function(j){
  
  rlej<-rle(x=diff(j$Index))
  position<-c(1,cumsum(rlej$lengths))
  timestack<-c()
  
  #if only one position
  if(length(position)==1){
    timestack[1]<-0
    
  } else {
    
    #time difference between first and last point in the cell
      for(x in 2:length(position)){
        timed<-j[position[x-1]:position[x],]
        timestack[x-1]<-difftime(max(timed$timestamp),min(timed$timestamp),units="days")
    }
  }

  out<-data.frame(Sum=sum(timestack))
  return(out)
}

#loop through all animals
sptracks<-lapply(sxy,function(x){
  k<-SpatialPointsDataFrame(cbind(x$x,x$y),data=x,proj4string=CRS("+proj=longlat +datum=WGS84"))
  
  #get the cells that points are in
  k$cellvalue<-raster::extract(rforaging,k)
  
  #for each cell value, perform time function
  sptrack<-as.data.frame(k) %>% group_by(Animal,cellvalue) %>% do(timecell(.))
  return(sptrack)}
)

sptracks<-rbind_all(sptracks)

#get sum across animals
cellvals<-sptracks %>% group_by(cellvalue) %>% summarize(Sum=sum(Sum))

#is there a better way to loop through?
rforaging[]<-NA
for(x in 1:nrow(cellvals)){
  rforaging[cellvals$cellvalue[x]]<-cellvals$Sum[x]
}
s[[3]]<-rforaging
names(s)[[3]]<-"Area-restricted Search"
}
```

```{r}

plot(s,nr=1,addfun=function(){map(add=T)})

#extract krill values
ksp<-data.frame(ksp,raster::extract(s,y=ksp))
mksp<-melt(ksp,measure.vars=names(s))

#write to file
ggplot(as.data.frame(mksp)) + geom_point(aes(y=value,x=log(C1KRIcatchKG),col=Unit))  + theme_bw() + labs(y="Residency Time (Days)",x="Krill Catch (Log Metric Tonnes)") + labs(col="Behavior") + facet_wrap(~variable)

ggplot(as.data.frame(mksp)) + geom_point(aes(y=value,x=log(C1KRIcatchKG),col=Unit),size=3)  + theme_bw() + labs(y="Residency Time (Days)",x="Krill Catch (Log Metric Tonnes)") + labs(col="Behavior") + facet_wrap(~variable,ncol=1) + scale_color_brewer(palette="Accent")
ggsave("Figures/ResidencyTime_ccamlr.jpeg",height=7,width=7,dpi=300)

mdf<-as.data.frame(mksp) 

camdf<-mdf %>% group_by(Unit,variable) %>% summarize(s=sum(value,na.rm=T)) %>% arrange(s) %>% ungroup() %>% mutate(Behavior=variable) %>% dplyr::select(-variable) %>%  dcast(...~Behavior,value.var="s") %>% arrange(All.Behavior)
write.csv(camdf,"Figures/CCAMLR_days.csv")
```

```{r}
temp <- get_map(location=bbox(r),source="google",zoom=5,maptype="satellite",color = "bw",scale = 2)

ggmap(temp) + geom_path(data=mxy,aes(x=x,y=y,col=phistate,group=paste(Animal,Track)),size=.5) + geom_point(data=krill,aes(x=GridMidpointDegreeLon,y=GridMidpointHalfDegreeLat,size=C1KRIcatchKG/1000)) + scale_fill_continuous(low="blue",high="red") + labs(fill="Days", size="Krill Catch (Metric Tons)") + scale_size_continuous(range=c(1,5)) + labs(col="Behavior") + labs(x="",y="") + theme(axis.text.x=element_blank(),axis.text.y=element_blank())
ggsave("Figures/TravelingResidenceTime_Krill.jpeg",height=9,width=7)
```

```{r}
ggmap(temp) + geom_tile(data=data.frame(rasterToPoints(r)),aes(x=x,y=y,fill=layer)) + geom_point(data=krill,aes(x=GridMidpointDegreeLon,y=GridMidpointHalfDegreeLat,size=C1KRIcatchKG/1000)) + scale_fill_continuous(low="blue",high="red") + labs(fill="Days", size="Krill Catch (Metric Tons)") + scale_size_continuous(range=c(1,5)) + labs(x="",y="") + theme(axis.text.x=element_blank(),axis.text.y=element_blank())
ggsave("Figures/ResidencyDays_Krill.jpeg",height=6,width=7,dpi=300)
```

```{r}
ggmap(temp) + geom_tile(data=data.frame(rasterToPoints(rforaging)),aes(x=x,y=y,fill=layer)) + geom_point(data=krill,aes(x=GridMidpointDegreeLon,y=GridMidpointHalfDegreeLat,size=C1KRIcatchKG/1000)) + scale_fill_continuous(low="blue",high="red",limits=c(0,150)) + labs(fill="Days", size="Krill Catch (Metric Tons)") + scale_size_continuous(range=c(1,5)) + ggtitle("Area Restricted Search") + labs(x="",y="") + theme(axis.text.x=element_blank(),axis.text.y=element_blank())
ggsave("Figures/ARSResidencyDays_Krill.jpeg",height=6,width=7,dpi=300)
```

```{r}
ggmap(temp) + geom_tile(data=data.frame(rasterToPoints(rtraveling)),aes(x=x,y=y,fill=layer)) + geom_point(data=krill,aes(x=GridMidpointDegreeLon,y=GridMidpointHalfDegreeLat,size=C1KRIcatchKG/1000)) + scale_fill_continuous(low="blue",high="red",limits=c(0,150)) + labs(fill="Days", size="Krill Catch (Metric Tons)") + scale_size_continuous(range=c(1,5)) + ggtitle("Traveling") + labs(x="",y="") + theme(axis.text.x=element_blank(),axis.text.y=element_blank())
ggsave("Figures/TravelingResidencyDays_Krill.jpeg",height=6,width=7,dpi=300)
```

```{r}
#check out the memory sizing
lsos()
#notify end of run
#size of the run
gc()
save.image("SingleSpecies.RData")
```