---
title: "Antarctic Whale Project: Single Species"
author: "Ben Weinstein"
date: "May 5th, 2016"
output: 
  html_document:
    keep_md: True
    numbered_sections: True
    toc: true
    theme: spacelab
---

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width = 10,fig.height = 5)
library(boot)
library(raster)
library(reshape2)
library(ggplot2)
library(MASS)
library(ggmap)
library(dplyr)
library(chron)
library(gridExtra)
library(stringr)
library(rjags)
library(move)

#New model to be run, flag turned off if just updating.
newModel<-T

#Source Mountains function, thanks to Sam Urmy http://www.oceanographerschoice.com/2010/10/fractal-landscapes-in-r-part-two/
source("Mountains.R")
```

```{r,eval=F}
#Load in data
load("SingleSpecies.RData")
jags$recompile()
newModel<-F
```

#Abstract

We are currently ignoring any observation process or time regularization.

```{r}
mdat<-move(x="InputData/Antarctic Humpback overlap with krill fisheries -2376078525850487209.csv")

#Create a oceandepth in km 
mdat$OceanDepth<-mdat$ETOPO1.Elevation*-1 / 1000
tm<-melt(timeLag(mdat,units="mins"))
paste("Mean time lag is:",round(mean(tm$value),2)/60,"hours")

#Plot
mdat_df <- as(mdat, "data.frame")
pp<-c(mean(mdat_df$location.long),mean(mdat_df$location.lat))
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite")
ggmap(m)+geom_path(data=mdat_df, aes(x=location.long, y=location.lat),col='red') + scale_color_discrete(guide=F)
```

#Correlated random walk

*Process Model*

$$ d_{t} \sim T*d_{t-1} + Normal(0,\Sigma)$$
$$ x_t = x_{t-1} + d_{t} $$

## Parameters

$$\theta = \text{Mean turning angle}$$
$$\gamma = \text{Move persistence} $$

###Behavioral States

$$ Behavior_1 = \text{traveling}$$
$$ Behavior_2 = \text{foraging}$$

$$ \alpha_{1,1} = \text{Probability of remaining traveling when traveling}$$
$$\alpha_{2,1} = \text{Probability of switching from feeding to traveling}$$

$$\begin{matrix}
  \alpha_{1,1} & 1-\alpha_{1,1} \\
  \alpha_{2,1} & 1-\alpha_{2,1} \\
\end{matrix}
$$

###Environment

Behavioral states are a function of local environmental conditions. The first environmental condition is ocean depth. I then build a function for preferential foraging in shallow waters.

It generally follows the form, conditional on behavior at t -1:

$$Behavior_t \sim Multinomial([\phi_{traveling},\phi_{foraging}])$$
$$logit(\phi_{traveling}) = \alpha_{Behavior_{t-1}} + \beta_1 * Ocean_{y[t-1,]}$$
$$logit(\phi_{foraging}) = \alpha_{Behavior_{t-1}} + \beta_2 * Ocean_{y[t-1,]}$$

#Model Fitting

```{r,eval=T}
#source model
source("Bayesian/RW.R")

#print model
print.noquote(readLines("Bayesian/RW.R"))

#prior shape
R <- diag(c(1,1))
data=list(y=cbind(mdat_df$location.long,mdat_df$location.lat),steps=nrow(mdat_df),R=R,ocean=mdat_df$OceanDepth)

if(newModel){
  jags<-jags.model(file = "Bayesian/RW.jags",data=data,n.chains=2)
  }

system.time(update(jags,1000))
p<-coda.samples(model=jags,n.iter=1000,thin=10,variable.names=c("theta","gamma","alpha","beta"))
```

##Chains
```{r,fig.height=5}
plot(p)
```

##Posteriors

```{r}
#bind chains
out<-rbind(p[[1]],p[[2]])
pc<-melt(out)
colnames(pc)<-c("Draw","par","value")

#Plot
ggplot(data=pc,aes(x=value)) + geom_histogram() + facet_wrap(~par,scales='free',ncol=4) + theme_bw()  + ggtitle("Estimated parameters, true values in red")

#As matrices
sv<-data.frame(str_match(pc$par,"(\\w+)\\[(\\d+)]"))[,-1]
colnames(sv)<-c("parname","index")

#format to combine
pc<-data.frame(pc,sv)
ggplot(pc,aes(x=index,y=value)) + geom_violin() + facet_wrap(~parname,scales="free")
```

Compare to priors

```{r}
#add comparison to priors
todraw<-table(pc$par)[1]

pc$Estimate<-"Posterior"

#add priors
#alpha
a1prior<-data.frame(Draw=1:todraw,par='alpha[1]',value=runif(todraw,-5,5),parname='alpha',index=1,Estimate="Prior")

a2prior<-data.frame(Draw=1:todraw,par='alpha[2]',value=runif(todraw,-5,5),parname='alpha',index=2,Estimate="Prior")

beta1prior<-data.frame(Draw=1:todraw,par='beta[1]',value=rnorm(todraw,0,0.0001),parname='beta',index=1,Estimate="Prior")

beta2prior<-data.frame(Draw=1:todraw,par='beta[2]',value=rnorm(todraw,0,0.0001),parname='beta',index=2,Estimate="Prior")

gamma1prior<-data.frame(Draw=1:todraw,par='gamma[1]',value=rbeta(todraw,5,2),parname='gamma',index=1,Estimate="Prior")

gamma2prior<-data.frame(Draw=1:todraw,par='gamma[2]',value=rbeta(todraw,2,5),parname='gamma',index=2,Estimate="Prior")

prs<-rbind_all(list(a1prior,a2prior,beta1prior,beta2prior,gamma1prior,gamma2prior))

prs$index<-as.factor(prs$index)

allp<-rbind_all(list(pc,prs))

ggplot(allp,aes(x=index,y=value,fill=Estimate)) + geom_violin() + facet_wrap(~parname,scale="free")

```

##Prediction - environmental function

```{r,fig.height=3,fig.width=3.5}
dplot<-function(x,a1,beta){

  y<-inv.logit(a1+beta*x)
  d11<-data.frame(x,y,State='Traveling',k='Traveling')
  
  y<-1-inv.logit((a1)+beta*x)
  d12<-data.frame(x,y,State='Traveling',k='Feeding')
  
  y<-inv.logit(a1+beta*x)
  d21<-data.frame(x,y,State='Feeding',k='Traveling')
  
  y<-1-inv.logit((a1)+beta*x)
  d22<-data.frame(x,y,State='Feeding',k='Feeding')
  
  d<-rbind_all(list(d11,d12,d21,d22))
  return(d)
  }
```

```{r}
alphap<-cbind(pc %>% filter(par=='alpha[1]') %>% .$value,pc %>% filter(par=='alpha[2]') %>% .$value)

betap<-cbind(pc %>% filter(par=='beta[1]') %>% .$value,pc %>% filter(par=='beta[2]') %>% .$value)

postplot<-list()

for(j in 1:nrow(alphap)){
  postplot[[j]]<-dplot(x=mdat$OceanDepth,a1=alphap[j,],beta=betap[j,])
}

postplot<-rbind_all(postplot)

dsum<-postplot %>% group_by(x,State,k) %>% summarize(y_mean=mean(y),lower=quantile(y,0.05),upper=quantile(y,0.95))

ggplot(dsum,aes(x=x,y=y_mean,ymin=upper,ymax=lower,col=k,fill=k)) + geom_point()  + geom_ribbon() + facet_wrap(~State) + theme_bw()
```

##Spatial Predicion

```{r,eval=T}
pr<-coda.samples(model=jags,n.iter=2000,n.thin=5,variable.names=c("d"))
ps<-rbind(pr[[1]],pr[[2]])
ps<-melt(ps)

#extract positions
sv<-data.frame(str_match(ps$Var2,"(\\w+)\\[(\\d+),(\\d+)]"))[,-1]
colnames(sv)<-c("par","step","xy")

#format to combine
ps<-data.frame(ps,sv)
#relevel
levels(ps$xy)<-c("x","y")

#Just keep mean, upper and lower
toplot<-dcast(ps,Var1+step~xy,value.var = "value")
sumplot<-group_by(toplot,step) %>% summarize(x_mean=mean(x),y_mean=mean(y))

#arrange by step
sumplot$step<-as.numeric(as.character(sumplot$step))
sumplot<-sumplot %>% arrange(step)

ggplot(sumplot,aes(x=x_mean,y=y_mean)) + geom_path(size=2,alpha=0.5) + theme_bw() + geom_path(data=mdat_df,aes(x=location.long,y=location.lat),col='red') + ggtitle("Predicted path (black) compared to true path (red)") 
```

Point Density

```{r,eval=T}
out<-rbind(pr[[1]],pr[[2]])
pc<-melt(out)
colnames(pc)<-c("Draw","param","value")

sv<-data.frame(str_match(pc$par,"(\\w+)\\[(\\d+),(\\d+)]"))[,-1]
colnames(sv)<-c("par","step","xy")

#format to combine
pc<-data.frame(pc,sv)
#relevel
levels(pc$xy)<-c("x","y")
toplot<-dcast(pc,Draw+step~xy,value.var = "value")

#arrange by step
toplot$step<-as.numeric(as.character(toplot$step))
toplot<-toplot %>% arrange(step)

ggplot(data=toplot,aes(x=x,y=y,col=step)) + geom_point() + theme_bw() + geom_path(data=mdat_df,aes(x=location.long,y=location.lat),col='red')
```

#Behavioral Prediction

## Point Cloud
```{r}
#needs to be same number of draws as spatial above
behav<-coda.samples(model=jags,n.iter=2000,n.thin=5,variable.names=c("phi"))

#bind chains
behav_chains<-rbind(behav[[1]],behav[[2]])
behav_chains<-melt(behav_chains)
colnames(behav_chains)<-c("Draw","par","value")

#As matrices
sv<-data.frame(str_match(behav_chains$par,"(\\w+)\\[(\\d+),(\\d+)]"))[,-1]
colnames(sv)<-c("parname","index")

#format to combine
behav_chains<-data.frame(behav_chains,sv)
colnames(behav_chains)<-c("Draw","par","Probability","parname","step","Behavior")

#We just need Feeding behavior
behav_chains<-behav_chains[behav_chains$Behavior == 2,]

#arrange by step
behav_chains$step<-as.numeric(as.character(behav_chains$step))
behav_chains<-behav_chains %>% arrange(step)

#Label Behaviors
behav_chains$Behavior<-as.factor(behav_chains$Behavior)
levels(behav_chains$Behavior)<-c("Traveling","Feeding")

#combine with spatial prediction
spp_b<-merge(toplot,behav_chains,by=c("Draw","step"))

#plot spatial
ggplot(data=spp_b,aes(x=x,y=y,col=Probability)) + geom_point() + theme_bw() + geom_path(data=mdat_df,aes(x=location.long,y=location.lat),col='red') + scale_color_continuous("Probability of Feeding",low='blue',high='red')
```

## Input data

```{r}
#get mean value
mean_phi<-behav_chains %>% group_by(step) %>% summarize(phi=mean(Probability))

#give mdat a step, column, this needs to be regularized.
mdat_df$Step<-1:nrow(mdat_df)

#merge with input
mdat_df<-merge(mdat_df,mean_phi,by.x="Step",by.y="step")

m <- get_map(location=pp,source="google",zoom=6,scale=2,maptype="satellite")
ggmap(m)+geom_path(data=mdat_df, aes(x=location.long, y=location.lat,col=phi),size=2.5) + scale_color_continuous("Probability of Feeding",low='blue',high='red')
```

##Autocorrelation in behavior

```{r}
ggplot(data=mdat_df,aes(x=as.numeric(Step),y=phi,col=OceanDepth)) + geom_line(aes(group=1),size=2.5) + labs("Probability of Feeding") + theme_bw() + scale_color_continuous("Ocean Depth (km)",low='light blue',high='black') + labs(x="Time Step",y="Liklihood of Feeding")
```

```{r,eval=T}
save.image("SingleSpecies.RData")
```
