---
title: "Antarctic Whale Project: Single Species"
author: "Ben Weinstein"
date: "`r Sys.time()`"
output: 
  html_document:
    keep_md: yes
    number_sections: yes
    theme: spacelab
    toc: yes
    pandoc_args: [
    "-M2GB", "+RTS", "-K64m", "-RTS"
    ]
---

```{r,warning=FALSE,message=FALSE,echo=FALSE}
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width = 11,fig.height = 5,cache=F)
library(boot)
library(raster)
library(reshape2)
library(ggplot2)
library(survival)
library(MASS)
library(ggmap)
library(dplyr)
library(chron)
library(gridExtra)
library(stringr)
library(R2jags)
library(move)

#New model to be run, flag turned off if just updating.
newModel<-F
```

```{r,eval=T}
#Load in data
load("SingleSpecies.RData")
#jags$recompile()
newModel<-F
```

```{r}
#memory function

# improved list of objects
.ls.objects <- function (pos = 1, pattern, order.by,
                        decreasing=FALSE, head=FALSE, n=5) {
    napply <- function(names, fn) sapply(names, function(x)
                                         fn(get(x, pos = pos)))
    names <- ls(pos = pos, pattern = pattern)
    obj.class <- napply(names, function(x) as.character(class(x))[1])
    obj.mode <- napply(names, mode)
    obj.type <- ifelse(is.na(obj.class), obj.mode, obj.class)
    obj.prettysize <- napply(names, function(x) {
                           capture.output(format(utils::object.size(x), units = "auto")) })
    obj.size <- napply(names, object.size)
    obj.dim <- t(napply(names, function(x)
                        as.numeric(dim(x))[1:2]))
    vec <- is.na(obj.dim)[, 1] & (obj.type != "function")
    obj.dim[vec, 1] <- napply(names, length)[vec]
    out <- data.frame(obj.type, obj.size, obj.prettysize, obj.dim)
    names(out) <- c("Type", "Size", "PrettySize", "Rows", "Columns")
    if (!missing(order.by))
        out <- out[order(out[[order.by]], decreasing=decreasing), ]
    if (head)
        out <- head(out, n)
    out
}

# shorthand
lsos <- function(..., n=10) {
    .ls.objects(..., order.by="Size", decreasing=TRUE, head=TRUE, n=n)
}
```

```{r}
#raster layers
#ice
ice<-raster("InputData/ETOPO1_Ice_g_geotiff.tif")
ice<-round(ice*-1/1000,3)
rice<-aggregate(ice,fact=0.1/res(ice))

rm(ice)

#coast
distc<-raster("InputData/DistancetoCoast.tif")
rice[rice<0]<-NA
distc[is.na(rice)]<-NA

#read data
mdat<-read.csv("InputData/Antarctic Humpback overlap with krill fisheries -3085052751565071394.csv")

##need to look into making into a movebank object.

#name env columns
#Create a oceandepth in km 
mdat$ocean<-round(mdat$ETOPO1.Elevation * -1 /1000,3)
mdat$coast<-round(mdat$NASA.Distance.to.Coast,1)
mdat$SST<-mdat$MODIS.Ocean.Aqua.OceanColor.4km.Monthly.Daytime.SST
mdat$ChloroA<-mdat$MODIS.Ocean.Aqua.OceanColor.4km.Monthly.Chlorophyll.A

#standardize column names to match the simulation
#Create an animal tag.
mxy <- as(mdat, "data.frame")
mxy$Animal<-mxy$individual.local.identifier
mxy$x<-mxy$location.long
mxy$y<-mxy$location.lat

#remove minke whales
minke<-c("131117","131118","131120","154184")
mxy<-mxy[!mxy$individual.local.identifier %in% minke,]

#grab set of animals

#mxy<-mxy[mxy$Animal %in% c("112699","121207","131151"),]

#empty coordinates
mxy<-mxy[!is.na(mxy$x),]

#crop by extent
d<-SpatialPointsDataFrame(cbind(mxy$x,mxy$y),data=mxy,proj4string=CRS("+proj=longlat +datum=WGS84"))

d<-crop(d,y=extent(c(-75,-50,-70,-57.5)))
#give it a geospatial layer


mxy<-as.data.frame(d)

#set datestamp
mxy$timestamp<-as.POSIXct(mxy$timestamp,format="%Y-%m-%d %H:%M:%S.000")

#month and year columns
mxy$Month<-months(mxy$timestamp)
mxy$Year<-years(mxy$timestamp)

#Only austral sping and summer
mxy<-mxy[mxy$Month %in% month.name[1:7],]

#remove empty timestamps
mxy<-mxy[!is.na(mxy$timestamp),]

#remove duplicates
mxy<-mxy[!duplicated(data.frame(mxy$timestamp,mxy$Animal)),]

```

#Descriptive Statistics
```{r}
#no distance to coast,elevation
mxy<-mxy[!mxy$ETOPO1.Elevation>0,]
```

```{r}
#hold copy of orig
oxy<-mxy

#remake spatial object
d<-SpatialPointsDataFrame(cbind(mxy$x,mxy$y),data=mxy,proj4string=CRS("+proj=longlat +datum=WGS84"))

#Plot
pp<-c(mean(mxy$x),mean(mxy$y))
m <- get_map(location=pp,source="google",zoom=4,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal)),size=.5) + scale_color_discrete(guide="none") + theme_inset()
```

##By Month

```{r,fig.height=12}
mxy$Month<-factor(mxy$Month,levels=month.name)
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal)),size=.5)  + facet_wrap(~Month) + scale_color_discrete(guide="none")
```

##Distance

```{r,fig.height=3,eval=F}
moved<-move(x=mxy$x,y=mxy$y,time=as.POSIXct(mxy$timestamp),proj=CRS("+proj=longlat +ellps=WGS84"),animal=mxy$Animal,data=as.data.frame(mxy),sensor=mxy$Animal,removeDuplicatedTimestamps=T)
dstat<-melt(move::distanceSummary(moved))

dstat<-droplevels(dstat[dstat$variable %in% c("TravDist","AverDist","FarthDist"),])
levels(dstat$variable)<-c("Total Distance","Average Distance Between Points","Farthest Distance Between Points")
ggplot(dstat,aes(value/1000)) + geom_histogram() + facet_wrap(~variable,scales="free") + labs(x="Kilometers")
```

##Time 
```{r,fig.height=3,eval=F}
tstat<-move::timeSummary(moved)
tstat<-melt(tstat,measure.vars=colnames(tstat[[1]]))
tstat<-droplevels(tstat[tstat$variable %in% c("Duration","AverDur"),])
levels(tstat$variable)<-c("Total Time (Hours)","Average Time Between Points (Hours)")
ggplot(tstat,aes(value)) + geom_histogram() + facet_wrap(~variable,scales="free")
```

##Velocity
```{r,fig.height=3,eval=F}
vstat<-move::speedSummary(moved)
vstat<-melt(vstat,measure.vars=colnames(vstat[[1]]))
levels(vstat$variable)<-c("Average Speed (m/s)","Variance in Speed (m/s)","Max Speed (m/s)")
ggplot(vstat,aes(value)) + geom_histogram() + facet_wrap(~variable,scales="free")
```

##Angles

```{r,fig.height=2,eval=F}
astat<-move::angleSummary(moved)
astat<-melt(astat,measure.vars=colnames(astat[[1]]))
astat<-droplevels(astat[astat$variable %in% "AverAzimuth",])
levels(astat$variable)<-"Turning Angle (degrees)"
ggplot(astat,aes(value)) + geom_histogram() + facet_wrap(~variable,scales="free")

#remove move object
rm(moved)
```

#Correlated random walk

*Process Model*

$$ d_{t} \sim T*d_{t-1} + Normal(0,\Sigma)$$
$$ x_t = x_{t-1} + d_{t} $$

## Parameters

For each individual:

$$\theta = \text{Mean turning angle}$$
$$\gamma = \text{Move persistence} $$

For both behaviors process variance is:
$$ \sigma_{latitude} = 0.1$$
$$ \sigma_{longitude} = 0.1$$

##Behavioral States

$$ \text{For each individual i}$$
$$ Behavior_1 = \text{traveling}$$
$$ Behavior_2 = \text{foraging}$$

$$ \alpha_{i,1,1} = \text{Probability of remaining traveling when traveling}$$
$$\alpha_{i,2,1} = \text{Probability of switching from Foraging to traveling}$$

$$\begin{matrix}
  \alpha_{i,1,1} & 1-\alpha_{i,1,1} \\
  \alpha_{i,2,1} & 1-\alpha_{i,2,1} \\
\end{matrix}
$$

##Environment

Behavioral states are a function of local environmental conditions. The first environmental condition is ocean depth. I then build a function for preferential foraging in shallow waters.

It generally follows the form, conditional on behavior at t -1:

$$Behavior_t \sim Multinomial([\phi_{traveling},\phi_{foraging}])$$

With the probability of switching states:

$$logit(\phi_{traveling}) = \alpha_{Behavior_{t-1}} + \beta_{Month,1} * Ocean_{y[t,]} + \beta_{Month,2} * Coast_{y[t,]}$$

$$logit(\phi_{foraging}) = \alpha_{Behavior_{t-1}} $$

Following Bestley in preferring to describe the switch into feeding, but no estimating the resumption of traveling.

The effect of the environment is temporally variable such that

$$ \beta_{Month,2} \sim ~ Normal(\beta_{\mu},\beta_\tau)$$


```{r,fig.height=3,fig.width=3.5}
dplot<-function(a1,beta=c(0,0),beta2=c(0,0),x=0,coast=0){

  #Transition from traveling to traveling
  y<-inv.logit(a1[1]+beta[1]*x + beta2[1]*coast)
  d11<-data.frame(x,coast,y,State="Traveling",Begin="Traveling")
  
  #transition from Traveling to Foraging
  y<-1-inv.logit(a1[1]+beta[1]*x+ beta2[1]*coast)
  d12<-data.frame(x,coast,y,State="Foraging",Begin="Traveling")

  y<-inv.logit(a1[2]+beta[2]*x+ beta2[2]*coast)
  d21<-data.frame(x,coast,y,State="Traveling",Begin="Foraging")
  
  y<-1-inv.logit(a1[2]+beta[2]*x+ beta2[2]*coast)
  d22<-data.frame(x,coast,y,State="Foraging",Begin="Foraging")
  
  d<-rbind_all(list(d11,d12,d21,d22))
  return(d)
}
```

##Continious tracks

The transmitter will often go dark for 10 to 12 hours, due to weather, right in the middle of an otherwise good track. The model requires regular intervals to estimate the turning angles and temporal autocorrelation. As a track hits one of these walls, call it the end of a track, and begin a new track once the weather improves. We can remove any micro-tracks that are less than three days.
Specify a duration, calculate the number of tracks and the number of removed points. Iteratively.

```{r}
##Time is the beginning of the first point.
step_length=6

sxy<-split(mxy,mxy$Animal)

#time diff function
timed<-function(d,step_length){
  d$j[1]<-0
  for (x in 2:nrow(d)){
    d$j[x]<-as.numeric(difftime(as.POSIXct(d$timestamp[x]),as.POSIXct(d$timestamp[x-1]),units="mins"))/(step_length*60)
  }
  
  #Split out track endings
  ends<-c(1,which(d$j>1),nrow(d))

  for(w in 2:length(ends)){
    d[ends[w-1]:ends[w],"Track"]<-w-1
  }
  
  #remove tracks that are shorter than three days
  track_time<-d %>% group_by(Track) %>% summarize(mt=difftime(max(as.POSIXct(timestamp)),min(as.POSIXct(timestamp)),units="hours")) %>% filter(mt>=24) %>% .$Track
  
  d<-d[d$Track %in% track_time,]
  
  #renumber the tracks
  d$Track<-as.numeric(as.factor(d$Track))
  return(d)
  }

sxy<-lapply(sxy,timed,step_length=6)

#Format matrices for jags
mxy<-rbind_all(sxy)

######recode whales
#mxy$Animal<-as.numeric(as.factor(mxy$Animal))

sxy<-split(mxy,list(mxy$Animal,mxy$Track),drop=TRUE)

sxy<-lapply(sxy,function(x){
#How many observations in each step length segment
x$Step<-as.numeric(cut(as.POSIXct(x$timestamp),"6 hours"))
return(x)
})

mxy<-rbind_all(sxy)

#refactor animal
mxy$Animal<-as.numeric(as.factor(mxy$Animal))
```

```{r}
#total number of steps per track/animal
steps_all<-mxy %>% group_by(Animal,Track) %>% summarize(n=length(unique(Step)))

# give each step a label
mxy<-mxy %>% group_by(Animal,Track,Step) %>% mutate(jStep=1:n())

#Cast time array
j<-acast(mxy,Animal~Track~Step~jStep,value.var="j")

#how many observations per individual in each Step
mxy$Step<-factor(mxy$Step,levels=1:max(steps_all$n))
idx<-melt(table(mxy$Animal,mxy$Track,mxy$Step))
colnames(idx)<-c("Animal","Track","Step","jStep")
idx<-acast(data=idx,Animal~Track~Step)

#make ocean a matrix -> MEAN VALUE -> will this yield a jags error on empty cells?
oc<-acast(mxy,Animal~Track~Step,value.var="ocean",fun.aggregate = mean)

#make coast a matrix -> MEAN VALUE
coast<-acast(mxy,Animal~Track~Step,value.var="coast",fun.aggregate = mean)

#month array
mxy$MonthF<-as.numeric(factor(mxy$Month,levels=month.name))

MonthA<-acast(mxy,Animal~Track~Step,value.var="MonthF",fun.aggregate = min)
MonthA[!is.finite(MonthA)]<-NA

#Individuals
ind=length(unique(mxy$Animal))

#tracks per indivudal
tracks<-mxy %>% group_by(Animal) %>% summarize(tracks=length(unique(Track))) %>% .$tracks

#steps per track
steps<-acast(steps_all,Animal~Track,value.var="n")

#obs array
obs<-melt(mxy,measure.vars=c("x","y"))
obs<-acast(obs,Animal~Track~Step~jStep~variable)
```

How did the filter change the extent of tracks?

```{r}
#Plot
mxy<-mxy %>% arrange(Animal,Track,Step,jStep)

pp<-c(mean(mxy$x),mean(mxy$y))
m <- get_map(location=pp,source="google",zoom=5,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=as.factor(Animal),group=paste(Animal,Track)),size=.5) + scale_color_discrete(guide="none") + theme_inset()
```

```{r,fig.height=15}
ggplot() +  geom_path(data=oxy,col='blue',aes(x=x,y=y)) +  geom_path(data=mxy,col='red',aes(x=x,y=y),size=2,alpha=0.7)  + facet_wrap(~individual.local.identifier,scales="free",ncol=4)
```

```{r,fig.height=3,fig.width=5}
ggplot(data=steps_all,aes(x=n*step_length/24)) + geom_histogram() + labs(x="Days") + ggtitle("Track Length")
ggplot(data=steps_all,aes(x=Track)) + geom_histogram() + labs(x="Subtracks per Animal")
```

```{r,child="Bayesian/MultiSpecies.R",eval=T}
```

```{r,eval=T}
#source jags file
source("Bayesian/MultiSpecies.R")

#prior cov shape
R <- diag(c(1,1))
data=list(argos=obs,steps=steps,R=R,ocean=oc,coast=coast,ind=ind,j=j,idx=idx,tracks=tracks,Month=MonthA,Months=max(MonthA,na.rm=T))

#paramters to track
pt<-c("theta","gamma","phi","alpha_mu","beta_mu","beta2_mu","beta","beta2")

if(newModel){
  system.time(jagM<-jags.parallel(model.file = "Bayesian/Multi_RW.jags",data=data,n.chains=2,parameters.to.save=pt,n.iter=50000,n.burnin=48000,n.thin=20,DIC=FALSE))
}

```

```{r,eval=T}
#update function
recompile(jagM)
jagM <- update(jagM,n.iter=5000,n.burnin=0,n.thin=5) 
```

##Chains
```{r,eval=T}

#check memory size
lsos()

#delete jags objects

rm(data)
rm(argos)
rm(obs)
rm(j)
gc()

#bind chains
pc<-melt(jagM$BUGSoutput$sims.array)

rm(jagM)
gc()

colnames(pc)<-c("Draw","chain","par","value")

#extract parameter name
pc$parameter<-data.frame(str_match(pc$par,"(\\w+)"))[,-1]

#Extract index
splitpc<-split(pc,pc$parameter)

#single index
splitpc[c("alpha_mu","beta_mu","beta2_mu","gamma","theta")]<-lapply(splitpc[c("alpha_mu","beta_mu","beta2_mu","gamma","theta")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc<-data.frame(x,Behavior=sv)
    return(pc)
})

## double index
splitpc[c("beta","beta2")]<-lapply(splitpc[c("beta","beta2")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+)]"))[,3:4]      
    colnames(sv)<-c("MonthF","Behavior")
    setp<-data.frame(x,sv)
    #get month name
    mindex<-mxy %>% ungroup() %>% select(MonthF,Month) %>% distinct()
    setp<-merge(setp,mindex,by="MonthF")
    return(setp)
})

#Three index
splitpc[c("phi")]<-lapply(splitpc[c("phi")],function(x){
#As matrices
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+),(\\d+)]"))[,3:6]
    colnames(sv)<-c("Animal","Track","step","Behavior")
    pc<-data.frame(x,sv)
})

#bind all matrices back together
pc<-rbind_all(splitpc)
rm(splitpc)


```

```{r,fig.height=20}
#plot all but phi and state
ggplot(pc[!pc$parameter %in% c("phi","state"),],aes(x=Draw,y=value)) + facet_wrap(~par,scales="free",ncol=3) + geom_line(aes(col=as.factor(chain))) + labs(col="Chain")
```


```{r,eval=F}
#write.csv
#write.csv(pc,"OutData/pars.csv")
#pc<-read.csv("OutData/pars.csv")
pc$Behavior<-as.factor(pc$Behavior)
```

```{r,fig.height=20}
#Plot
ggplot(data=pc[!pc$parameter %in% c("phi","state"),],aes(x=value)) + geom_histogram() + facet_wrap(~par,scales='free',ncol=3) + theme_bw()  + ggtitle("Estimated parameters")
```

###Compare to priors

```{r}
#add comparison to priors
todraw<-table(pc$par)[1]
pc$Estimate<-"Posterior"

#add priors
#alpha
a1prior<-data.frame(Draw=1:todraw,par='alpha[1]',value=rnorm(todraw,0,1.67),parameter='alpha_mu',Behavior=1,Estimate="Prior")

a2prior<-data.frame(Draw=1:todraw,par='alpha[2]',value=rnorm(todraw,0,1.67),parameter='alpha_mu',Behavior=2,Estimate="Prior")

beta1prior<-data.frame(Draw=1:todraw,par='beta[1]',value=rnorm(todraw,0,1.67),parameter='beta_mu',Behavior=1,Estimate="Prior")

beta21prior<-data.frame(Draw=1:todraw,par='beta2[1]',value=rnorm(todraw,0,1.67),parameter='beta2_mu',Behavior=1,Estimate="Prior")

gamma1prior<-data.frame(Draw=1:todraw,par='gamma[1]',value=rbeta(todraw,1,10),parameter='gamma',Behavior=1,Estimate="Prior")

gamma2prior<-data.frame(Draw=1:todraw,par='gamma[2]',value=rbeta(todraw,1.5,5),parameter='gamma',Behavior=2,Estimate="Prior")

prs<-rbind_all(list(a1prior,a2prior,beta1prior,beta21prior,gamma1prior,gamma2prior))

prs$Behavior<-as.factor(prs$Behavior)

allp<-rbind_all(list(pc[!pc$parameter %in% c("phi","state"),],prs))

ggplot(allp,aes(x=Behavior,y=value,fill=Estimate)) + geom_violin() + facet_wrap(~parameter,scale="free")
rm(allp)
```

## Parameter Summary
```{r,fig.height=10,fig.width=15}
psum<-pc %>% filter(!parameter %in% c("phi","state")) %>% group_by(parameter,par) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

as.data.frame(psum)

write.csv(psum,"Figures/Table1.csv")
ggplot(psum,aes(x=par,y=mean,ymax=upper,ymin=lower)) + geom_linerange() + geom_point() + facet_wrap(~parameter,scales="free",ncol=2)
```

#Behavior and environment

##Hierarchical 

### Ocean Depth
```{r}
#get the posterior estimates of the env function
postplot<-pc %>% filter(parameter %in% c('beta_mu','alpha_mu')) %>% dcast(.,Draw+chain~par) %>% group_by(Draw,chain) %>% do(dplot(a1=c(.$'alpha_mu[1]',.$'alpha_mu[2]'),beta=c(.$'beta_mu[1]',.$'beta_mu[2]'),x=unique(mxy$ocean),coast=0)) %>% group_by(x=x,Begin,State) %>% summarize(mean=mean(y),lower=quantile(y,0.05),upper=quantile(y,0.95))

ggplot(postplot) + geom_ribbon(aes(x=x,y=mean,ymin=upper,ymax=lower,fill=State),alpha=0.5) + theme_bw() + labs(col="Transition",x="Ocean Depth (km)",y="Probability",fill="Transition") + facet_wrap(~Begin)+ xlim(0,quantile(mxy$ocean,0.75))

#mean env estimate
ggplot(postplot) + geom_ribbon(aes(x=x,y=mean,ymin=upper,ymax=lower,fill=State),alpha=0.5) + theme_bw() + labs(col="Transition",x="Ocean Depth (km)",y="Probability",fill="Transition") + facet_wrap(~Begin) 

ggplot(postplot[postplot$Begin %in% "Traveling" & postplot$State %in% "Foraging",]) + geom_ribbon(aes(x=x,y=mean,ymin=upper,ymax=lower),alpha=0.2) + theme_bw() + labs(col="Transition",x="Ocean Depth (km)",y="Probability of switching to foraging state",fill="Transition") + geom_line(aes(x=x,y=mean)) 
ggsave("Figures/Ocean.jpeg",dpi=300,height=3,width=4)
```

### Distance to Coast
```{r}
#get the posterior estimates of the env function
postplot<-pc %>% filter(parameter %in% c('beta2_mu','alpha_mu')) %>% dcast(.,Draw+chain~par) %>% group_by(Draw,chain) %>% do(dplot(a1=c(.$'alpha_mu[1]',.$'alpha_mu[2]'),beta2=c(.$'beta2_mu[1]',.$'beta2_mu[2]'),coast=unique(mxy$coast))) %>% group_by(coast,Begin,State) %>% summarize(mean=mean(y),lower=quantile(y,0.05),upper=quantile(y,0.95))

#mean env estimate
ggplot(postplot) + geom_ribbon(aes(x=coast,y=mean,ymin=upper,ymax=lower,fill=State),alpha=0.5) + theme_bw() + labs(col="Transition",x="Distance to Coast (km)",y="Probability",fill="Transition") + facet_wrap(~Begin) + xlim(0,quantile(mxy$coast,0.75))

ggplot(postplot) + geom_ribbon(aes(x=coast,y=mean,ymin=upper,ymax=lower,fill=State),alpha=0.5) + theme_bw() + labs(col="Transition",x="Distance to Coast (km)",y="Probability",fill="Transition") + facet_wrap(~Begin) 

ggplot(postplot[postplot$Begin %in% "Traveling" & postplot$State %in% "Foraging",]) + geom_ribbon(aes(x=coast,y=mean,ymin=upper,ymax=lower),alpha=0.2) + theme_bw() + labs(col="Transition",x="Distance to Coast (km)",y="Probability of switching to foraging state",fill="Transition") + geom_line(aes(x=coast,y=mean)) + xlim(0,quantile(mxy$coast,0.75))

ggsave("Figures/Distance_Coast.jpeg",dpi=300,height=3,width=4)

```

###Interaction

```{r}
coastx<-seq(0,quantile(mxy$coast,0.98),quantile(mxy$coast,0.98)/20)
oceanx<-seq(0,quantile(mxy$ocean,0.98),quantile(mxy$ocean,0.98)/20)
allx<-expand.grid(coastx,oceanx)
colnames(allx)<-c("coast","ocean")
allx$ID<-1:nrow(allx)
allx$alpha<-filter(pc,par %in% 'alpha_mu[1]') %>% group_by(parameter) %>% summarize(m=mean(value)) %>% .$m
allx$beta1<-filter(pc,par %in% 'beta_mu[1]') %>% group_by(parameter) %>% summarize(m=mean(value)) %>% .$m
allx$beta2<-filter(pc,par %in% 'beta2_mu[1]') %>% group_by(parameter) %>% summarize(m=mean(value)) %>% .$m

traj<-function(ocean,coast,alpha,beta1,beta2){
  p<-1-inv.logit(alpha + beta1*ocean + beta2*coast)
  data.frame(ocean,coast,phi=p)
}

intplot<-allx %>% group_by(ID) %>% do(traj(.$ocean,.$coast,.$alpha,.$beta1,.$beta2))

ggplot(intplot,aes(x=ocean,y=coast,fill=phi)) + geom_tile() + theme_bw() + labs(x="Ocean Depth (km)",y="Distance to coast (km)") + scale_fill_gradient(low="blue",high="red") + ggtitle("Mean Probability of Foraging when Traveling")
```

## By Month

### Depth

```{r}
#get the posterior estimates of the env function
postplot_alpha<-pc %>% filter(parameter %in% c('alpha_mu')) %>% do(dcast(.,Draw+chain~par))
postplot_beta<-pc %>% filter(parameter %in% c('beta')) %>% dcast(.,Month+Draw+chain~parameter+Behavior)
postall<-merge(postplot_alpha,postplot_beta,by=c("Draw","chain"))

allplot<-postall %>% group_by(Month,Draw,chain) %>% do(dplot(a1=c(.$'alpha_mu[1]',.$'alpha_mu[2]'),beta=c(.$'beta_1',.$'beta_2'),x=unique(mxy$ocean),coast=0)) %>% group_by(x=x,Month,Begin,State) %>% summarize(mean=mean(y),lower=quantile(y,0.05),upper=quantile(y,0.95))

allplot$Month<-factor(allplot$Month,levels=month.name)

ggplot(allplot) + geom_ribbon(aes(x=x,y=mean,ymin=upper,ymax=lower,fill=Month),alpha=0.5) + theme_bw() + labs(col="Transition",x="Ocean Depth (km)",y="Probability",fill="Transition") + facet_grid(State~Begin)+ xlim(0,quantile(mxy$ocean,0.75)) 

#mean env estimate
ggplot(allplot) + geom_ribbon(aes(x=x,y=mean,ymin=upper,ymax=lower,fill=Month),alpha=0.5) + theme_bw() + labs(col="Month",x="Ocean Depth (km)",y="Probability",fill="Transition") + facet_grid(State~Begin) 
```

Just the probability of feeding when traveling.

```{r}
#mean env estimate
ggplot(allplot[allplot$Begin=="Traveling" & allplot$State == "Foraging",]) + geom_ribbon(aes(x=x,y=mean,ymin=upper,ymax=lower,fill=Month),alpha=0.5,size=1) + theme_bw() + labs(col="Transition",x="Ocean Depth (km)",y="Probability of switching to foraging state",fill="Month") + scale_fill_brewer(palette="YlOrRd") + facet_wrap(~Month)
```

Just mean estimate.

```{r}
ggplot(allplot[allplot$Begin=="Traveling" & allplot$State == "Foraging",]) + geom_line(aes(x=x,y=mean,ymin=upper,ymax=lower,col=Month),alpha=0.5,size=1) + theme_bw() + labs(col="Transition",x="Ocean Depth (km)",y="Probability of Feeding",fill="Month") + xlim(0,quantile(mxy$ocean,0.75))

#mean env estimate
ggplot(allplot[allplot$Begin=="Traveling" & allplot$State == "Foraging",]) + geom_line(aes(x=x,y=mean,ymin=upper,ymax=lower,col=Month),alpha=0.5,size=1) + theme_bw() + labs(col="Month",x="Ocean Depth (km)",y="Probability of switching to foraging state",fill="Month") + scale_color_brewer(palette="YlOrRd")
```

### Coast

```{r}
#get the posterior estimates of the env function
postplot_alpha<-pc %>% filter(parameter %in% c('alpha_mu')) %>% do(dcast(.,Draw+chain~par))
postplot_beta<-pc %>% filter(parameter %in% c('beta2')) %>% dcast(.,Month+Draw+chain~parameter+Behavior)
postall<-merge(postplot_alpha,postplot_beta,by=c("Draw","chain"))

allplot<-postall %>% group_by(Month,Draw,chain) %>% do(dplot(a1=c(.$'alpha_mu[1]',.$'alpha_mu[2]'),beta2=c(.$'beta2_1',.$'beta2_2'),coast=unique(mxy$coast))) %>% group_by(coast,Month,Begin,State) %>% summarize(mean=mean(y),lower=quantile(y,0.05),upper=quantile(y,0.95))

allplot$Month<-factor(allplot$Month,levels=month.name)

ggplot(allplot) + geom_ribbon(aes(x=coast,y=mean,ymin=upper,ymax=lower,fill=Month),alpha=0.5) + theme_bw() + labs(col="Transition",x="Distance to Coast (km)",y="Probability",fill="Month") + facet_grid(State~Begin)+ xlim(0,quantile(mxy$coast,0.75)) + scale_fill_brewer(palette="YlOrRd")

#mean env estimate
ggplot(allplot) + geom_ribbon(aes(x=coast,y=mean,ymin=upper,ymax=lower,fill=Month),alpha=0.5) + theme_bw() + labs(col="Transition",x="Distance to Coast (km)",y="Probability",fill="Month") + facet_grid(State~Begin) + scale_fill_brewer(palette="YlOrRd")
```

Zooming in on the top right plot.
```{r}
ggplot(allplot[allplot$Begin=="Traveling" & allplot$State == "Foraging",]) + geom_ribbon(aes(x=coast,y=mean,ymin=upper,ymax=lower,fill=Month),alpha=0.5,size=1) + theme_bw() + labs(col="Transition",x="Distance to coast (km)",y="Probability of Feeding",fill="Month") + xlim(0,quantile(mxy$coast,0.75)) + scale_fill_brewer(palette="YlOrRd")

#mean env estimate
ggplot(allplot[allplot$Begin=="Traveling" & allplot$State == "Foraging",]) + geom_ribbon(aes(x=coast,y=mean,ymin=upper,ymax=lower,fill=Month),alpha=0.5,size=1) + theme_bw() + labs(col="Transition",x="Distance to coast (km)",y="Probability of switching to foraging state",fill="Month") + scale_fill_brewer(palette="YlOrRd")

ggplot(allplot[allplot$Begin=="Traveling" & allplot$State == "Foraging",]) + geom_ribbon(aes(x=coast,y=mean,ymin=upper,ymax=lower,fill=Month),alpha=0.5,size=1) + theme_bw() + labs(col="Transition",x="Distance to coast (km)",y="Probability of switching to foraging state",fill="Month") + scale_fill_brewer(palette="YlOrRd") + facet_wrap(~Month)

```

Just mean estimate.

```{r}
ggplot(allplot[allplot$Begin=="Traveling" & allplot$State == "Foraging",]) + geom_line(aes(x=coast,y=mean,ymin=upper,ymax=lower,col=Month),alpha=0.5,size=1) + theme_bw() + labs(col="Transition",x="Ocean Depth (km)",y="Probability of Feeding",fill="Month") + xlim(0,quantile(mxy$coast,0.75))+ scale_color_brewer(palette="YlOrRd")

#mean env estimate
ggplot(allplot[allplot$Begin=="Traveling" & allplot$State == "Foraging",]) + geom_line(aes(x=coast,y=mean,ymin=upper,ymax=lower,col=Month),alpha=0.5,size=1) + theme_bw() + labs(col="Month",x="Ocean Depth (km)",y="Probability of switching to foraging state",fill="Month") + scale_color_brewer(palette="YlOrRd")
```

#Behavioral Prediction

```{r}
#We just need Foraging behavior
behav_chains<-pc[pc$parameter %in% "phi" & pc$Behavior==2,]

#arrange by time - capitalize
behav_chains$Step<-as.numeric(as.character(behav_chains$step))
behav_chains<-behav_chains %>% arrange(Step)

#Label Behaviors
behav_chains$Behavior<-as.factor(behav_chains$Behavior)
levels(behav_chains$Behavior)<-c("Traveling","Foraging")

mean_phi<-behav_chains %>% group_by(Animal,Track,Step) %>% summarize(phi=mean(value))

mxy<-merge(mxy,mean_phi,by=c("Step","Track","Animal"))

#ensure order for plotting
mxy<-mxy %>% arrange(Animal,Track,Step,jStep)

rm(behav_chains)
```

##Spatial Prediction

```{r}
m <- get_map(location=pp,source="google",zoom=4,scale=2,maptype="satellite",color="bw")
ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=phi,group=paste(Animal,Track)),size=.5)   + scale_color_continuous(low='blue',high='red',limits=c(0,1)) + labs(col="Probability of Foraging")  + theme_inset() + geom_polygon(data=pols[[2]],aes(x=long,y=lat,group=id),fill="whitesmoke")
```

### Per Animal
```{r,eval=F}
#ggmap(m)+geom_path(data=mxy, aes(x=x, y=y,col=phi,group=Track),size=.5)   + scale_color_continuous(low='blue',high='red',limits=c(0,1)) + labs(col="Probability of Foraging") + facet_wrap(~individual.local.identifier) + theme_inset() 
```

##Log Odds of Foraging

### Ocean Depth

```{r,fig.height=20}
#phi and ocean plot
ggplot(data=mxy,aes(x=ocean,y=logit(phi),col=Month)) + geom_point() + labs(x="Ocean Depth",y="Estimated Log Odds of Foraging") + facet_wrap(~individual.local.identifier,ncol=3,scales="free") + geom_smooth(method="lm")+ scale_color_brewer(palette="YlOrRd")
```

### Distance From Coast

```{r,fig.height=20}
#phi and ocean plot
ggplot(data=mxy,aes(x=coast,y=logit(phi),col=Month)) + geom_point() + labs(x="Distance to coast",y="Estimated Log Odds of Foraging") + facet_wrap(~individual.local.identifier,ncol=3,scales="free")+ geom_smooth(method="lm") + scale_color_brewer(palette="YlOrRd")
```

##Autocorrelation in behavior

```{r,fig.height=20,fig.width=13}
#create proper time stamp
ggplot(data=mxy,aes(x=as.POSIXct(timestamp),y=phi,col=ocean)) + geom_line(aes(group=Track),size=2.5) + labs("Probability of Foraging") + theme_bw() + scale_color_continuous(low='light blue',high='black') + labs(x="Time",y="Liklihood of Foraging") + facet_wrap(~individual.local.identifier,ncol=3,scales="free") + ylim(0,1)
```

##Behavioral description

##Location of Behavior

```{r}
msp<-SpatialPointsDataFrame(cbind(mxy$x,mxy$y),data=mxy[,c("x","y","Animal","phi","timestamp")],proj=CRS("+proj=longlat +datum=WGS84"))

r<-raster(msp,ext=extent(c(-75,-50,-70,-57.5)))
res(r)<-0.1

m<-rasterize(x=msp,y=r,field="phi")

#plotting
feedmap <- data.frame(rasterToPoints(m))

temp <- get_map(location=bbox(m),source="google",zoom=4,maptype="satellite",color = "bw",scale = 2)

#plot(m)

ggmap(temp) + labs(fill="Probabilty of Foraging")+ geom_tile(data=feedmap,aes(x=x, y=y,fill=layer),alpha=0.9) + theme_minimal() + scale_color_discrete(guide="none") + scale_fill_continuous(low="blue",high="red")
```

#Environmental Prediction - Probability of Foraging across time

```{r}
#get data frame of ice events
fns<-paste("InputData/SeaIce/",month.abb[1:7],"_2015","LatLong.csv",sep="")
pols<-lapply(fns,function(x){read.csv(x)})
names(pols)<-month.abb[1:7]
```

## Bathymetry

```{r}
meanpar<-pc %>% filter(parameter %in% c('beta_mu','alpha_mu'))  %>% group_by(par) %>% summarize(mean=mean(value)) %>% dcast(...~par)

#remove non-water areas
rice[rice<0]<-NA

#Traveling
travpar<-calc(rice,function(x){ inv.logit(1-(meanpar$`alpha_mu[1]` + meanpar$`beta_mu[1]` * x))})

plot(travpar)

#travdf<-data.frame(rasterToPoints(travpar))
  
#ggmap(temp) + labs(fill="Mean Probabilty of Foraging")+ geom_tile(data=travdf,aes(x=x, y=y,fill=layer),alpha=0.9) + scale_color_discrete(guide="none") + scale_fill_continuous(low="blue",high="red",limits=c(0,1)) + xlim(-75,-50) + ylim(-70,-57.5) 
```

## Distance to coast

```{r}
meanpar<-pc %>% filter(parameter %in% c('beta2_mu','alpha_mu'))  %>% group_by(par) %>% summarize(mean=mean(value)) %>% dcast(...~par)

#remove non-water areas
rice[rice<0]<-NA

#Traveling
travpar<-calc(distc,function(x){ 1-inv.logit(meanpar$`alpha_mu[1]` + meanpar$`beta2_mu[1]` * x)})
#travdf<-data.frame(rasterToPoints(travpar))

plot(travpar)
#ggmap(temp) + labs(fill="Mean Probabilty of Foraging")+ geom_tile(data=travdf,aes(x=x, y=y,fill=layer),alpha=0.9) + scale_color_discrete(guide="none") + scale_fill_continuous(low="blue",high="red") + xlim(-75,-50) + ylim(-70,-57.5)
```

##All variables

```{r}
meanpar<-pc %>% filter(parameter %in% c('beta_mu',"beta2_mu",'alpha_mu'))  %>% group_by(par) %>% summarize(mean=mean(value)) %>% dcast(...~par)

##Bathymetry
s<-stack(rice,distc)

#Traveling
travpar<-calc(s,function(x,y){ 1-inv.logit(meanpar$`alpha_mu[1]` + meanpar$`beta_mu[1]` * x[1] + meanpar$`beta2_mu[1]` * x[2])})

plot(travpar)

travdf<-data.frame(rasterToPoints(travpar))

#ggmap(temp) + labs(fill="Mean Probabilty of Foraging")+ geom_tile(data=travdf,aes(x=x, y=y,fill=layer),alpha=0.9)  + scale_color_discrete(guide="none") + scale_fill_continuous(low="blue",high="red")+theme(axis.text.x=element_blank(),axis.text.y=element_blank()) + xlim(-75,-50) + ylim(-70,-57.5) + labs(x="",y="")
```

# Overlap with Krill Fishery
```{r}
krill<-read.csv("InputData/CCAMLR_aggregated_catch_C1.csv")

ggmap(temp) + geom_tile(data=travdf,aes(x=x, y=y,fill=layer),alpha=0.7) + geom_point(data=krill,aes(x=GridMidpointDegreeLon,y=GridMidpointHalfDegreeLat,size=C1KRIcatchKG/1000)) + scale_fill_continuous(low="blue",high="red") + labs(fill="Probability of Foraging", size="Krill Catch (Metric Tons)") + scale_size_continuous(range=c(0,8)) 
```

```{r}
ksp<-SpatialPointsDataFrame(cbind(krill$GridMidpointDegreeLon,krill$GridMidpointHalfDegreeLat),krill)
#extract probability of feeding at krill midpoints
ksuit<-extract(x=travpar,y=ksp)
kdf<-data.frame(ksp,ksuit)

ggplot(kdf,aes(y=ksuit,x=log(C1KRIcatchKG))) + geom_point() + theme_bw() + labs(x="Log KG Tonnes of Krill Catch",y="Probability of switching from traveling to foraging") + geom_smooth(method="glm",method.args=list(family="binomial"))
ggsave("Figures/Krillcatch.jpeg",height=4,width=4,dpi=300)
```

## By Month

```{r}
#get the posterior estimates of the env function
postplot_alpha<-pc %>% filter(parameter %in% c('alpha_mu')) %>% do(dcast(.,Draw+chain~par))
postplot_beta<-pc %>% filter(parameter %in% c('beta',"beta2")) %>% dcast(.,Month+Draw+chain~parameter+Behavior)
postall<-merge(postplot_alpha,postplot_beta,by=c("Draw","chain"))

postall$Month<-factor(postall$Month,levels=month.name)
Months<-unique(postall$Month)

monthras<-list()
for (x in Months){

  monthpar<-postall %>% filter(Month==x)
  
  ##Bathymetry
  s<-stack(rice,distc)
  
  #Traveling
  travpar<-calc(s,function(x){ 1-inv.logit(mean(monthpar$`alpha_mu[1]`) + mean(monthpar$`beta_1` ) * x[1] + mean(monthpar$`beta2_1` * x[2]))})
  
  travdf<-data.frame(rasterToPoints(travpar))
  
  monthras[[x]]<-travpar
}
names(monthras)<-Months
```

```{r}
  #plot
for(x in 1:length(monthras)){  
    ggmap(temp) + labs(fill="Mean Probabilty of Foraging")+   geom_tile(data=data.frame(rasterToPoints(monthras[[x]])),aes(x=x, y=y,fill=layer),alpha=0.9)  + scale_color_discrete(guide="none") + scale_fill_continuous(low="blue",high="red",limits=c(0,1))+  ggtitle(names(monthras)[[x]])+theme(axis.text.x=element_blank(),axis.text.y=element_blank()) + xlim(-75,-50) + ylim(-70,-57.5) + labs(x="",y="") + geom_polygon(data=pols[[x]],aes(x=long,y=lat,group=group),fill="whitesmoke")
}
```

## Change in foraging areas

###Jan verus May

Red = Better Foraging in Jan
Blue = Better Foraging in May

```{r}
diffras<-monthras[[1]]-monthras[[5]]
diffdf<-as.data.frame(rasterToPoints(diffras))
ggmap(temp) + geom_tile(data=diffdf,aes(x=x, y=y,fill=layer)) + scale_fill_gradient2("",low='blue',midpoint=0,high='red')  + geom_polygon(data=pols[[2]],aes(x=long,y=lat,group=id),fill="whitesmoke")
ggsave("Figures/")
```

### Variance in monthly suitability

```{r}
plot(varArea<-calc(stack(monthras),sd))
```

### Mean suitability

```{r}
plot(varArea<-calc(stack(monthras),mean))
```

## Monthly Overlap with Krill Fishery

```{r}
ksp<-SpatialPointsDataFrame(cbind(krill$GridMidpointDegreeLon,krill$GridMidpointHalfDegreeLat),krill)
#extract probability of feeding at krill midpoints
ksuit<-extract(x=stack(monthras),y=ksp)
kdf<-data.frame(ksp,ksuit)
mdf<-melt(kdf,measure.vars=names(monthras))

mdf$Month<-factor(mdf$variable,levels=month.name)

ggplot(mdf,aes(y=value,x=C1KRIcatchKG,col=Month)) + geom_point() + theme_bw() + labs(x="Log KG Tonnes of Krill Catch",y="Foraging Habitat Suitability") + geom_smooth(method="glm",method.args=list(family="binomial"))
```


```{r}
#check out the memory sizing
lsos()
#notify end of run
#size of the run
gc()
save.image("SingleSpecies.RData")
```
